\documentclass{article}

\newcommand{\BCRM}{BCRM}

\begin{document}

\section{A common language for production networks}

Let there be $J$ industries. Index a using industry by $i$ and a supplying
industry by $j$, and let $\Omega\in\mathbb{R}^{J\times J}_{+}$ use the
user-by-supplier convention. If the spectral radius satisfies
$\rho(\Omega)<1$, then $I-\Omega$ is invertible and its Neumann series
converges:

\begin{equation}
  \mathcal{L}=(I-\Omega)^{-1}=I+\Omega+\Omega^2+\cdots
  \label{eq:leontief}
\end{equation}

Here $I$ is the $J\times J$ identity matrix. The term $I$ is own exposure,
$\Omega$ is direct supplier exposure, and $(\Omega^n)_{ij}$ collects length-$n$
paths from user $i$ to upstream supplier $j$. Equation~\eqref{eq:leontief} is
an accounting map. The infinite-series equality requires $\rho(\Omega)<1$;
mere invertibility of $I-\Omega$ would not by itself justify that series.

If $\beta\in\mathbb{R}^{J}$ is a column vector of final-consumption shares, the
corresponding Domar weights are defined by

\begin{equation}
  \lambda^{\prime}=\beta^{\prime}(I-\Omega)^{-1}.
  \label{eq:domar}
\end{equation}

These weights measure gross activity, not welfare and not an impulse response.

\section{A quantitative multisector DSGE: \BCRM{} (2009)}

\subsection{Question and model architecture}

Sector $j$ produces gross output with capital, labor, and materials. The source
model maintains constant returns in these three inputs:

\begin{equation}
  y_{jt}=k_{jt}^{\alpha_j}(z_t n_{jt})^{\nu_j}H_{jt}^{\gamma_j},
  \qquad \alpha_j+\nu_j+\gamma_j=1,
  \label{eq:bcrm-production}
\end{equation}

The restriction $\alpha_j+\nu_j+\gamma_j=1$ is a source-stated maintained
assumption. The material composite is defined by

\begin{equation}
  H_{jt}=\left(\sum_{i=1}^{J}\zeta_{ij}
  h_{ji,t}^{(\theta-1)/\theta}\right)^{\theta/(\theta-1)}.
  \label{eq:bcrm-material}
\end{equation}

Here $h_{ji,t}$ is user $j$'s purchase from supplier $i$, and $\zeta_{ij}$ is
the source's supplier weight in user $j$'s material bundle.

\end{document}
