\chapter{Work Packages and Delivery Plan}

This appendix breaks the project into concrete work packages.

\section*{WP1: Document and code indexing}
\begin{itemize}
  \item Build LaTeX extraction for chapters, labels, theorem-like blocks, equations, and references.
  \item Build code inventory for functions, classes, tests, and relevant symbols.
  \item Produce local caches and unified search.
  \item Deliver first retrieval benchmark.
\end{itemize}

\section*{WP2: Verification backends}
\begin{itemize}
  \item Add SymPy and SymEngine first-line symbolic checks.
  \item Add mpmath numeric probing.
  \item Add SageMath and Maxima fallback wrappers.
  \item Add Z3-based consistency checks for encodable cases.
  \item Normalize outputs into structured result records.
\end{itemize}

\section*{WP3: MCP interface}
\begin{itemize}
  \item Implement a local Python MCP server.
  \item Expose a small tool surface with structured schemas.
  \item Add project-local configuration for Claude Code.
  \item Create examples and smoke tests.
\end{itemize}

\section*{WP4: Audit and consistency workflows}
\begin{itemize}
  \item Implement claim auditing and gap localization.
  \item Implement code--document consistency checks.
  \item Produce batch reports and benchmark outputs.
\end{itemize}

\section*{WP5: Code generation support}
\begin{itemize}
  \item Define mathematically constrained generation tasks.
  \item Add acceptance-test harnesses based on formulas and examples.
  \item Benchmark generated code against baseline workflows.
\end{itemize}

\section*{WP6: Adoption and documentation}
\begin{itemize}
  \item Write user-facing workflow guides.
  \item Create sample benchmark tasks and example reports.
  \item Train group members on interactive and batch usage patterns.
\end{itemize}
