\section{State-space preliminaries}
The first part fixes notation for a long-document retrieval stress test. Several labels use similar covariance and innovation language, so search should not rely on a unique phrase appearing only once.

\begin{equation}\label{eq:longdoc-transition-covariance}
P_{t|t-1} = T_t P_{t-1|t-1}T_t' + R_t Q_t R_t'.
\end{equation}

The transition covariance is only a prediction object. It does not define the smoothing gain or the score contribution.

\section{Innovation bookkeeping}
The innovation residual and innovation covariance are reused by several later formulas.

\begin{equation}\label{eq:longdoc-innovation-residual}
v_t = y_t - Z_t a_{t|t-1}.
\end{equation}

\begin{equation}\label{eq:longdoc-innovation-covariance}
S_t = Z_t P_{t|t-1}Z_t' + H_t.
\end{equation}

These equations are dependencies for likelihood derivatives. They should remain distinct from any smoothed-state identity.

\section{Likelihood derivative notes}
The local score contribution combines trace, quadratic, innovation residual, and innovation covariance derivative terms.

\begin{equation}\label{eq:longdoc-score-contribution}
\partial_i \ell_t
= -\frac{1}{2}\operatorname{tr}(S_t^{-1}\partial_i S_t)
+ \frac{1}{2}v_t' S_t^{-1}(\partial_i S_t)S_t^{-1}v_t
- (\partial_i v_t)'S_t^{-1}v_t.
\end{equation}

This local score contribution is not a Hessian proof. Differentiating it requires additional derivative recursions that are not included here.

\section{Smoothing recursions}
The smoothing pass introduces a nearby but different gain matrix. It shares state-space notation with the likelihood section, which makes it useful for provenance stress testing.

\begin{equation}\label{eq:longdoc-smoothing-gain}
J_t = P_{t|t}T_{t+1}'P_{t+1|t}^{-1}.
\end{equation}

The gain equation mentions covariance matrices and time indices, but it is not evidence for the score contribution above.

\section{Observed information reminder}
The observed information block uses second derivatives of the likelihood. The score contribution above is one dependency, not the complete observed information identity.

\begin{equation}\label{eq:longdoc-observed-information}
\mathcal I_{ij} = -\partial_{ij}^2 \ell(\theta).
\end{equation}

A provenance-aware tool should retrieve the score contribution when asked about trace residual derivative terms, not the smoothing gain or the generic observed information reminder.
