\documentclass[12pt,a4paper,openany]{book}

\input{preamble}

\title{\Huge\textbf{MathDevMCP}\\[10pt]
  \Large Final Release Report for an Industrial Mathematical Development Agent}
\author{Chak Wong and collaborators}
\date{April 29, 2026}

\begin{document}

\frontmatter
\maketitle
\tableofcontents

\mainmatter

\part{Release Claim}

\chapter{Executive Summary}

Mathematical teams already know the failure modes that matter here. A note says a
likelihood uses a log determinant and a linear solve, while the implementation
has been refactored twice under deadline. A proof sketch depends on
positive-definiteness or invertibility, but the assumption is buried in another
section or carried only in a senior colleague's memory. A new agent can sound
confident about the mathematics, but the answer is hard to trust because it does
not show where the claim came from or where the verification boundary ends.

MathDevMCP is built for those moments. It gives colleagues a local-first way to
connect claims to the implementation, surface support and mismatches, and avoid
making unsupported mathematical statements. Instead of asking a reviewer to trust
an agent's paraphrase, it turns local LaTeX documents, labels, code, parser
signals, typed obligations, backend checks, and release gates into structured
evidence packets that a human can inspect, challenge, and reuse.

The product promise is practical rather than grand. MathDevMCP helps a colleague
find the right equation and its nearby assumptions, detect that code no longer
carries a documented solve-like operation, decompose a derivation into typed
obligations with explicit backend evidence or abstention, check scoped symbolic
equalities with deterministic backends, and hand another reviewer a
document-grounded implementation brief.
It is most useful in review loops where unsupported claims are expensive and
where a generic assistant is too willing to smooth over missing provenance.

This is why the release claim is deliberately narrow and evidence based.
MathDevMCP does not claim to prove arbitrary mathematical papers correct. Its
value is that it preserves the distinction between retrieval, typed obligation
diagnostics, scoped equality checks, mismatch, abstention, and deterministic
certification in machine-readable reports. In practice, that means colleagues can
move faster without pretending that a search hit, a parser extraction, a typed
shape diagnostic, or a nearby symbolic pattern is already a proof.

The current full release profile is configured to require benchmark evidence,
parser policy, governance validation, release-corpus metadata, LeanDojo backend
environment evidence, LaTeXML validation, and a private-corpus manifest. The
private-corpus evidence used for this report is an external sanitized corpus
generated outside the repository. Its purpose is to validate the privacy and
release-gate machinery without committing private department documents. A real
department manifest can be substituted through the same environment variable.

\begin{proposalbox}
\textbf{Release status.} With an external sanitized private manifest supplied,
the full profile has no blocking findings. Generated report snippets may record
a dirty working tree while the report is being edited; the post-commit release
readiness gate should be rerun before a release tag is cut.
\end{proposalbox}

\section{Why This Tool Exists}

MathDevMCP exists because mathematical review work breaks down in predictable
ways. Derivations drift from code, hidden assumptions disappear into nearby
sections, and ordinary assistants can retrieve plausible context without showing
whether the implementation carries the required operation or whether the
derivation remains uncertified. The tool is useful precisely because it keeps
labels, file paths, line spans, code comparisons, typed diagnostics, and
certification boundaries visible. It does not replace expert judgment; it gives
expert judgment a better starting point.

\chapter{MCP Conversations That Sell The Tool}

The most natural way to use MathDevMCP is not as a long list of commands. It is
as a conversation with an agent that has a small set of disciplined tools. The
colleague asks the same question they would ask in a hallway or code review:
"Does this implementation still match the equation in the note?" The difference
is that the agent can answer by calling bounded MCP tools, citing labels and
files, and preserving the verification boundary.

This chapter is intentionally early in the report because it is the adoption
story. A colleague does not need to learn every command before seeing value.
They need to learn the shape of a good prompt, the kind of tool call it
triggers, and the kind of answer that should come back.

\section{The Conversation Pattern}

A productive MCP conversation usually has five turns:

\begin{enumerate}[label=\arabic*.]
  \item ask the agent to locate the mathematical claim in a local root;
  \item ask for the surrounding assumptions or notation;
  \item ask the agent to compare the claim with a concrete code file;
  \item ask whether the remaining derivation is certified, diagnostic, or
        unverified;
  \item ask for a short handoff note that another reviewer can reproduce.
\end{enumerate}

The important behavior is not that the agent sounds confident. The important
behavior is that it chooses tools such as \texttt{search\_latex},
\texttt{latex\_label\_lookup}, \texttt{audit\_implementation\_label},
\texttt{audit\_derivation\_v2\_label}, \texttt{typed\_obligation\_label},
\texttt{check\_equality}, and \texttt{implementation\_brief} before making a
mathematical claim.

\section{A Realistic Review Transcript}

\begin{lstlisting}[basicstyle=\ttfamily\scriptsize,caption={MCP conversation for a document-code review}]
User:
  In /path/to/project, check whether the Kalman likelihood in the notes
  still matches src/kalman_filter.py. I care about logdet, solve, and the
  quadratic form. Keep the proof boundary explicit.

Agent:
  I will first find the likely label, then compare that label to the code.

Tool: search_latex
  root=/path/to/project
  query="Kalman likelihood log determinant solve"
  limit=3

Tool result:
  label=eq:dept-state-space-likelihood
  file=doc_department_state_space.tex
  lines=...

Agent:
  I found label eq:dept-state-space-likelihood. I will inspect nearby context
  before comparing the implementation.

Tool: latex_label_lookup
  root=/path/to/project
  label=eq:dept-state-space-likelihood

Tool result:
  status=found
  context includes the likelihood block and nearby state-space assumptions

Tool: audit_implementation_label
  root=/path/to/project
  label=eq:dept-state-space-likelihood
  code=/path/to/project/src/kalman_filter.py
  required_terms=["logdet", "solve", "quadratic_form"]
  before=1, after=1

Tool result:
  status=mismatch
  matched_terms=["logdet"]
  missing_terms=["solve", "quadratic_form"]

Agent:
  The checked code path visibly carries logdet, but the configured review
  terms solve and quadratic_form are missing. This is not a theorem that the
  code is wrong. It is a document-grounded drift signal. The next review step
  is to inspect whether the quadratic term is implemented through an equivalent
  factorization or whether the solve-like operation was dropped.

User:
  Can you also tell me whether the derivation itself is certified?

Tool: audit_derivation_v2_label
  root=/path/to/project
  label=eq:dept-state-space-likelihood
  summary_only=true

Tool result:
  status=unverified
  high_priority_actions=[
    "state or verify invertibility",
    "check log-det domain",
    "check conformability",
    "check linear-solve residual"
  ]

Agent:
  The derivation audit is diagnostic, not a certificate. I would write the
  review note as: the label was found, the implementation comparison reports
  a missing solve/quadratic-form signal, and the derivation remains unverified
  until the matrix assumptions and solve residual are reviewed.
\end{lstlisting}

\section{Prompt Patterns That Trigger Tools}

\begin{longtable}{p{0.38\textwidth}p{0.25\textwidth}p{0.27\textwidth}}
\toprule
Prompt shape & MCP tool & Why it helps \\
\midrule
"Find the equation or assumption for ..." & \path{search_latex} &
Returns labels and provenance instead of memory. \\
"Show the context around label ..." & \path{latex_label_lookup} &
Surfaces nearby assumptions and notation. \\
"Compare this label to this code file ..." & \path{audit_implementation_label} &
Turns a vague drift worry into matched and missing terms. \\
"Check whether this derivation is certified." &
\path{audit_derivation_v2_label} &
Keeps the distinction between proof, diagnostic evidence, and abstention. \\
"Give me a handoff note for review." & \path{implementation_brief} &
Bundles search, context, comparison, and next action. \\
"Run the release-profile-analysis and tell me which claims are ready versus
which still need external evidence." &
\path{release_profile_analysis} &
Summarizes readiness across base, public, backend, LaTeXML, private-corpus,
and full release claims. \\
"Is this release ready?" & \path{release_readiness} &
Reports blockers and caveats instead of informal readiness claims. \\
\bottomrule
\end{longtable}

\section{When MCP Tools Trigger Instead of LLM-Only Reasoning}

A useful mental model is that the agent has two modes. In an LLM-only answer,
the agent reasons from the prompt, the chat history, and any text the user has
pasted into the conversation. That can be appropriate for ordinary explanation,
brainstorming, rewriting, or conceptual discussion. In an MCP-grounded answer,
the agent asks MathDevMCP for bounded evidence from the repository: labels,
file paths, line spans, code comparisons, derivation diagnostics, benchmark
results, or release-readiness reports.

The difference matters because MathDevMCP does not automatically run on every
mathematical word in a conversation. A client model still decides whether to
call a tool. The user can make the decision almost unavoidable by asking for
local evidence, naming a repository root, naming a label or code path, or
asking for a status that only a tool should produce. If the response contains
no tool name, no arguments, no label provenance, and no result fields, the user
should treat it as an LLM-only answer.

\begin{longtable}{p{0.30\textwidth}p{0.30\textwidth}p{0.30\textwidth}}
\toprule
Situation & LLM-only answer is reasonable & MCP-grounded answer is expected \\
\midrule
Conceptual explanation &
"Explain why Kalman likelihoods contain a log determinant." The answer can
explain the mathematics without inspecting local files. &
"In this repository, find the likelihood label and cite the file and line
span." This asks for local provenance, so \path{search_latex} should run. \\
Editing prose &
"Make this paragraph clearer." The user supplies the text and does not ask for
repository evidence. &
"Rewrite this paragraph after checking the nearby theorem assumptions." The
agent should retrieve context with \path{latex_label_lookup}. \\
Code review &
"Does this function look idiomatic?" This is a general programming judgment. &
"Compare label \path{eq:dept-state-space-likelihood} with
\path{src/kalman_filter.py} for \path{logdet} and \path{solve}." The agent
should call \path{audit_implementation_label}. \\
Derivation review &
"Is this derivation convincing?" Without a label or root, the agent may only
reason from pasted text. &
"Audit label \path{eq:dept-state-space-likelihood} and tell me whether the
result is certified, diagnostic, or unverified." This points to
\path{audit_derivation_v2_label}. \\
Release handoff &
"Are we ready to release?" This is ambiguous unless the repository and profile
are identified. &
"Run release readiness for this repository under the full profile and list
blockers and caveats." This should trigger \path{release_readiness}, and often
\path{benchmark_gate} or \path{doctor}. \\
\bottomrule
\end{longtable}

\subsection*{Signals That Should Cause Tool Use}

The following prompt features are strong trigger signals. They do not guarantee
tool use in every client, but they tell both the agent and the reviewer that an
LLM-only answer would be incomplete.

\begin{enumerate}[label=\arabic*.]
  \item The prompt names a local repository root, project directory, manifest,
        LaTeX file, or code file.
  \item The prompt asks the agent to find, cite, or inspect a label, theorem,
        assumption, equation, section, or paragraph in local documents.
  \item The prompt asks whether a document claim still matches a code path.
  \item The prompt gives required mathematical or implementation terms such as
        \path{logdet}, \path{solve}, \path{gradient}, \path{hamiltonian}, or
        \path{euler_residual}.
  \item The prompt asks for a status such as \path{consistent},
        \path{mismatch}, \path{unverified}, \path{ready},
        \path{ready_with_caveats}, or \path{not_ready}.
  \item The prompt asks for derivation certification, typed obligations,
        backend evidence, abstentions, or proof boundaries.
  \item The prompt asks for release evidence, benchmark gates, private corpus
        validation, governance policy, backend availability, or profile
        readiness.
  \item The prompt says "use MCP", "show the tool result", "cite labels and
        files", "do not answer from memory", or "keep the evidence boundary
        explicit".
\end{enumerate}

\subsection*{Cases That Look Like Tool Triggers But Usually Are Not}

Many plausible prompts sound technical but still invite an LLM-only response.
These are useful failure cases to show colleagues because they explain why
prompt shape matters. The improved prompt does not need to be verbose; it only
needs to specify the local artifact and the evidence expected.

\begin{longtable}{p{0.30\textwidth}p{0.28\textwidth}p{0.32\textwidth}}
\toprule
Prompt that may not trigger MCP & Why it may stay LLM-only & Better prompt \\
\midrule
"Explain the Kalman likelihood." &
This asks for general mathematical exposition, not local evidence. &
"In \path{/path/to/project}, find the Kalman likelihood label and summarize it
with file and line provenance." \\
"Does the filter implementation look right?" &
There is no code path, document label, or required term list. &
"Compare \path{eq:dept-state-space-likelihood} with
\path{src/kalman_filter.py}; check \path{logdet}, \path{solve}, and the
quadratic form." \\
"Can MathDevMCP prove this derivation?" &
The product audits bounded obligations; it does not promise unrestricted proof
repair. &
"Run \path{audit_derivation_v2_label} on this label and report whether the
result is certified, diagnostic, unverified, or blocked." \\
"The paper and code might be inconsistent. Thoughts?" &
The agent can speculate because no local target is supplied. &
"Search the LaTeX for the likelihood label, then compare the best label to
\path{src/likelihood.py} with required terms \path{logdet} and \path{solve}." \\
"Check the private corpus." &
The tool needs a manifest path or configured environment variable, and the
answer must respect redaction policy. &
"Validate this private manifest with MathDevMCP, do not print private paths,
and report release-gate blockers only." \\
"Find recent papers about Hamiltonian Monte Carlo." &
That is a literature or web task, not a local MathDevMCP repository task. &
"In this repository, find the HMC leapfrog label and compare it to the sampler
implementation for \path{gradient} and \path{hamiltonian}." \\
"Summarize the equation I pasted below." &
The user supplied the whole object, so the agent may summarize directly. &
"After summarizing the pasted equation, search the repository for matching
labels and cite any local match before making implementation claims." \\
"Run the release check." &
The release tool requires a root and, for full evidence, a profile and optional
private manifest configuration. &
"Run \path{release_readiness} for \path{/path/to/project} with profile
\path{full}, then list blockers, caveats, and missing backend evidence." \\
\bottomrule
\end{longtable}

\subsection*{How A Colleague Can Verify That The Tool Actually Ran}

The safest adoption practice is to ask the agent to expose the evidence trail.
A tool-grounded answer should show at least the tool name, important arguments,
and result fields. For example, a comparison answer should mention
\path{audit_implementation_label}, the label, the code path, matched terms, missing
terms, and the status. A derivation answer should mention
\path{audit_derivation_v2_label}, the parser route, obligations, backend
evidence or abstentions, and high-priority review actions. A release answer
should mention \path{release_readiness}, the profile, blockers, caveats, and
capability evidence.

If those artifacts are absent, the answer may still be useful, but it should be
treated as explanation rather than repository evidence. The most reliable
follow-up is direct: "Please rerun this using MathDevMCP tools, show the tool
calls and result statuses, and do not make claims that are not supported by the
tool output." That instruction converts an ambiguous prompt into an auditable
MCP workflow.

\section{Example Prompts and Expected Tool Calls}

This is the colleague-facing cheat sheet for proposal and adoption discussions.
The user does not need to name tools directly. They should describe the research
or review problem in ordinary language, and the agent should translate that
prompt into one or more narrow MCP calls before answering. If the agent answers
without first grounding the request in labels, file paths, or release evidence,
the colleague should ask it to rerun the task using MathDevMCP.

\begin{longtable}{p{0.34\textwidth}p{0.25\textwidth}p{0.31\textwidth}}
\toprule
Example prompt to the agent & Tools that should trigger & Evidence the answer should cite \\
\midrule
"Find where the model note defines the likelihood used by this filter." &
\path{search_latex} &
Candidate labels, source files, line spans, and short snippets rather than a
free-memory summary. \\
"Show me the assumptions around \path{eq:dept-state-space-likelihood} before
you judge the implementation." &
\path{latex_label_lookup} &
The labeled block plus nearby assumptions about dimensions, invertibility,
conditioning, and notation. \\
"Compare the likelihood equation with \path{src/kalman_filter.py}; I care
about \path{logdet}, \path{solve}, and the quadratic form." &
\path{audit_implementation_label} &
A status such as \path{consistent}, \path{mismatch}, or
\path{unverified}, with matched terms, missing terms, label provenance, and
code provenance. \\
"Can you check whether this derivation is actually certified, or only
diagnostic?" &
\path{audit_derivation_v2_label} &
Parser route, typed obligations, backend evidence when available, abstentions,
and high-priority human review actions. \\
"Before I edit the code, give me a compact implementation brief for this label
and file." &
\path{implementation_brief} &
Search result, extracted context, comparison result, caveats, and a short next
action that another developer can reproduce. \\
"Run the release-facing checks for this repository and tell me what blocks a
handoff." &
\path{benchmark_gate}, \path{doctor}, \path{release_readiness} &
Benchmark totals, expected abstentions, backend availability, release profile,
blockers, and caveats. \\
"Validate this private corpus manifest, but do not print private paths." &
\path{validate_release_corpus}, \path{release_readiness} &
Redacted corpus metadata, gate categories, blocked private paths, and the
profile-specific readiness result. \\
\bottomrule
\end{longtable}

The expected user habit is simple: ask for the mathematical task in natural
language, insist on the evidence packet, and treat the agent's prose as a
summary of tool output rather than as an independent authority. That habit is
what turns MathDevMCP from a command collection into a practical colleague
assistant.

The rest of the report shows this same pattern across the public case studies.
Each case gives the mathematical situation, the generated evidence, and a short
MCP-style transcript that a colleague can adapt directly in their own client.

\chapter{Product Scope}

MathDevMCP supports five base colleague-facing activities plus first-wave
productization surfaces for domain templates, proof packets, negative evidence,
claim support, status taxonomy, and backend readiness:

\begin{enumerate}[label=\arabic*.]
  \item locating mathematical context in LaTeX source trees;
  \item extracting labeled equation, theorem, assumption, and paragraph
        neighborhoods with provenance;
  \item comparing document claims with Python implementations;
  \item auditing derivation obligations with parser, typed, shape, numeric, and
        backend evidence;
  \item producing repeatable release evidence across base, backend, LaTeXML,
        private-corpus, and full profiles;
  \item turning recurring domain families into explicit diagnostic obligations;
  \item packaging proof state, negative evidence, and literature support without
        confusing those packets with mathematical proof;
  \item reporting local status vocabulary and Lean/Lake/LeanDojo readiness so an
        agent knows which next step is actually available.
\end{enumerate}

In colleague terms, this means the tool helps answer five recurring workflow
questions:

\begin{enumerate}[label=\alph*.]
  \item Where is the equation or assumption we are talking about?
  \item What nearby notation or assumptions does it depend on?
  \item Does the code visibly carry the documented structure?
  \item If a derivation is not certified, why not?
  \item Can another reviewer reproduce the evidence without trusting my memory?
  \item Which domain template, assumption manifest, convention registry, or
        proof packet should be used next?
\end{enumerate}

The May 2026 implementation changes are important because the tool now has
explicit vocabulary for product surfaces that were previously only plans:
source-local equation rows, assumption manifests, convention registries,
dependency graphs, matrix/operator IR diagnostics, proof and negative-evidence
packets, code-contract diagnostics, a small domain-template catalog, local claim
support packets, status/substatus taxonomy, and Lean readiness checks.

The product is intentionally local-first. Source documents and code remain on
the user's machine. Optional tools such as LaTeXML, Lean, LeanDojo, Sage, and
SymPy are detected at runtime and reported through structured diagnostics.
LeanDojo remains in the isolated \texttt{mathdevmcp-backends} environment so its
dependency stack does not destabilize the main working environment.

\section{What is included}

The release includes:

\begin{itemize}
  \item a LaTeX indexer with label, environment, context extraction, and
        source-local equation-row localization;
  \item code and document search utilities;
  \item code--document consistency checks;
  \item derivation-step checks with explicit \emph{unverified} and
        \emph{mismatch} outcomes;
  \item proof obligation, scoped equality, typed-obligation, and proof-audit
        workflows;
  \item typed and dimensional diagnostic workflows;
  \item status and substatus taxonomy for conservative workflow reports;
  \item assumption-manifest parsing and linting for label-level declarations;
  \item notation and sign-convention registry primitives;
  \item dependency-graph and impact-report primitives for labels, conventions,
        assumptions, packets, and code references;
  \item matrix/operator IR diagnostics for ordered products, inverses, log
        determinants, solves, traces, quadratic forms, Jacobians, and related
        unresolved constructs;
  \item AST operation extraction for mathematical implementation patterns;
  \item code-contract diagnostics for shape, dtype, batch-axis, finite-value,
        and XLA/custom-op boundaries;
  \item proof-packet and negative-evidence packet builders;
  \item a first-wave domain-template catalog for Kalman likelihood, CIP/SDF
        sign conventions, and HMC transform/Jacobian obligations;
  \item local claim-support packets for assumptions, empirical regularities,
        proposed extensions, citations, labels, and proof packets;
  \item direct Lean, Lake, and LeanDojo readiness diagnostics;
  \item parser benchmark and parser policy selection;
  \item release corpus manifest validation;
  \item private-corpus validation with redacted output;
  \item release-readiness profiles;
  \item MCP facade and FastMCP server wrappers;
  \item operator, deployment, release, security, and maintainer documentation.
\end{itemize}

\section{What is not claimed}

MathDevMCP does not replace expert mathematical review. A parser hit is not a
proof. A shape diagnostic is not a theorem. A Lean skeleton is not a certificate
until direct Lean checking accepts it without placeholders. A numeric diagnostic
can refute a false claim or suggest a failure mode, but it does not establish a
general theorem. This boundary is a product feature: it lets colleagues move
faster without laundering a plausible hint into a mathematical certificate.

\section{Review Workflow}

A normal assistant workflow tends to move from question to answer too quickly.
MathDevMCP changes the order of operations: find the labeled source context,
inspect nearby assumptions, compare the documented claim against code, audit the
certification boundary, and hand off a compact evidence packet. That ordering is
the main usability story of the tool because it makes another reviewer's local
reproduction path explicit.

\chapter{Release Profiles}

The release profiles are:

\begin{description}
  \item[base] Public benchmarks, parser policy, governance, release corpus, and
    normal doctor evidence.
  \item[backend] Base evidence plus isolated LeanDojo backend environment
    validation.
  \item[latexml] Base evidence plus strict LaTeXML parser validation.
  \item[private-corpus] Base evidence plus an external private or sanitized
    private manifest.
  \item[full] All of the above.
  \item[public] Product-surface release evidence: CI workflow presence,
    packaging metadata, MCP registry and README consistency, support-matrix
    coverage, documentation boundary language, quality-gate execution, and
    generated-evidence redaction.
\end{description}

The profile split is important because colleagues may use the base product
without every optional backend, while release managers can demand strict
evidence before a broader deployment. Optional evidence is not silently ignored:
strict profiles turn missing evidence into blocking findings.

The \texttt{full} profile is deliberately an internal/deployment profile. It
means that optional internal evidence sources are present; it is not by itself a
public industrial release claim. The \texttt{public} profile adds the product
surface checks needed before a public release statement: MCP documentation must
match the registry and FastMCP wrappers, the support matrix must cover every
release profile, CI must run the local gates, and default public error
envelopes must be structured and free of raw private paths.

\section{Full profile evidence}

\lstinputlisting[caption={Full profile release-readiness summary},basicstyle=\ttfamily\footnotesize]{generated/release_report/release-readiness-full-summary.txt}

\section{Doctor evidence}

\lstinputlisting[caption={Doctor capability summary},basicstyle=\ttfamily\footnotesize]{generated/release_report/doctor-summary.txt}

\part{Architecture}

\chapter{System Architecture}

MathDevMCP has three layers. The first layer is a Python implementation
substrate that owns parsing, indexing, consistency checks, proof-audit routing,
benchmarking, release policy, and diagnostics. The second layer is a CLI and MCP
facade that presents stable tools to users and agents. The third layer is a
workflow and documentation layer that explains how to combine tools for
colleague tasks.

The new product split is a core-platform layer plus optional domain packs. The
core platform owns contracts, statuses, provenance, equation localization,
assumption manifests, convention records, dependency graphs, matrix/operator
IR, proof packets, negative evidence, code contracts, claim support, and backend
readiness. Domain packs should stay declarative: they use those primitives to
generate review obligations for recurring mathematical families rather than
embedding a separate macro-finance application inside the core.

The architecture is intentionally plain. The project does not hide release logic
inside a server process or a notebook. A release manager can run the same Python
library functions through the CLI, through MCP wrappers, and through tests. This
is why the report can include command output instead of screenshots or manual
claims: the system is designed around reproducible JSON contracts.

\section{Implementation substrate}

The Python package is organized around small report-producing functions. Each
public report includes a metadata contract so that CLI, MCP, tests, and release
docs can depend on stable shapes. For example, release readiness returns a
\texttt{release\_readiness\_report}; parser benchmark returns
\texttt{parser\_backend\_result}; private corpus validation returns
\texttt{private\_corpus\_validation\_report}. This is the main compatibility
discipline for future maintainers.

The ownership map is practical:
\path{release_policy.py} decides blockers and caveats;
\path{release_corpus.py} owns public/private manifest assembly and path
redaction; \path{parser_policy.py} chooses whether parser output is strong
enough for proof-audit routing; \path{proof_audit_v2.py} carries the
verification-boundary language; \path{equation_locator.py} owns source-local
equation rows; \path{assumption_manifest.py}, \path{conventions.py}, and
\path{dependency_graph.py} own the diagnostic context graph;
\path{matrix_ir.py} owns the first matrix/operator IR route;
\path{proof_packet.py} and \path{negative_evidence.py} own durable review
packets; \path{domain_templates.py} owns the first governed template catalog;
\path{code_contracts.py} owns implementation-boundary diagnostics;
\path{claim_support.py} owns non-proof citation and claim support packets;
\path{lean_readiness.py} separates direct Lean, Lake, and LeanDojo readiness;
\path{mcp_facade.py} and \path{cli.py} expose library behavior to agents and
humans. A maintainer changing a release claim should know which one of those
files owns the claim before editing the report.

\section{CLI and MCP facades}

The CLI is the reproducible interface for humans and CI. The MCP facade wraps
the same library functions so an agent can call them interactively. The facade
does not grant unrestricted shell access. It exposes bounded tools such as
\path{search_latex}, \path{audit_implementation_label},
\path{audit_derivation_v2_label}, \path{benchmark_gate}, and
\path{release_readiness}. The newly exposed MCP surfaces are intentionally
narrow: \path{status_taxonomy} tells an agent which statuses and substatuses are
allowed, and \path{lean_readiness} reports direct Lean, Lake, and LeanDojo
availability separately.

Several May 2026 surfaces are CLI/library-first rather than MCP tools:
\path{proof-packet-label}, \path{negative-evidence-label},
\path{domain-templates}, \path{suggest-domain-templates},
\path{generate-template-obligations}, and \path{claim-support}. Keeping these
packet and catalog builders CLI-first for the initial release is deliberate.
They can produce larger artifacts, and they are easier to review in CI or as
files before becoming default conversational tools.

\section{Backend isolation}

Optional backends are treated as evidence engines. LaTeXML, Pandoc, Lean, Sage,
SymPy, and LeanDojo are detected by \texttt{doctor}. LeanDojo is checked through
the backend environment when configured. This avoids dependency conflicts in
the main Python environment while still allowing proof-search experiments and
Lean fixture validation.

This split matters operationally. Colleagues can use search, context extraction,
consistency checks, and release-corpus validation in the main environment. A
stricter release can then demand LaTeXML and LeanDojo evidence without forcing
everyday users to import every optional backend. The release profiles make that
distinction visible rather than relying on informal environment knowledge.

\chapter{Core Data Contracts}

Every user-facing workflow should be understood as a contract. A contract has a
schema version, a name, a status or result payload, and where appropriate
provenance. This is why downstream documentation, agents, and tests can
interpret outputs without scraping prose.

The contract discipline is also what makes MathDevMCP maintainable. The CLI can
pretty-print JSON, the MCP facade can return the same dictionary to an agent,
and the release report can include a shortened evidence file without each layer
inventing its own vocabulary. When a colleague sees \texttt{mismatch} in a
brief, that word means the same thing as it means in the benchmark gate and in
the tests.

\section{Status vocabulary}

The primary statuses are:

\begin{description}
  \item[verified] A deterministic backend accepted the exact scoped obligation
    under explicit assumptions.
  \item[consistent] The requested terms or structure were found in the checked
    context.
  \item[equivalent] A scoped derivation step matched exactly after normalization
    or backend checking.
  \item[unverified] Evidence exists, but it is not enough for certification.
  \item[mismatch] Required structure is missing or contradicted.
  \item[inconclusive] The tool lacks enough structure, parser support, or
    backend evidence to decide.
\end{description}

This vocabulary is central to the product. It prevents the agent from turning a
nearby equation or diagnostic hint into a categorical mathematical claim.

Substatuses make conservative results actionable without weakening that
boundary. Examples include
\texttt{unverified:missing\_assumption},
\texttt{unverified:missing\_shape},
\texttt{unverified:parser\_limit},
\texttt{unverified:unsupported\_noncommutative\_algebra},
\texttt{unverified:manual\_formalization\_required},
\texttt{mismatch:likely\_formula\_error},
\texttt{mismatch:normalization\_gap},
\texttt{inconclusive:source\_label\_missing},
\texttt{inconclusive:partial\_index\_only},
\texttt{inconclusive:backend\_unavailable},
\texttt{inconclusive:toolchain\_not\_ready}, and
\texttt{inconclusive:timeout}. The \texttt{status-taxonomy} CLI command and
\texttt{status\_taxonomy} MCP tool expose this vocabulary to agents.

\section{Equation localization}

The LaTeX index now carries equation-row diagnostics in addition to labeled
blocks. Each row reports source file, line span, extracted equation text, a
\texttt{localization\_status}, and uncertainty markers such as macro
non-expansion or preserved alignment markers. This matters for real monographs:
proof-audit and template suggestions should know whether they are looking at a
precisely localized equation row or a row with parser uncertainty.

The localization contract is still provenance evidence, not proof evidence. A
row can make the source easier to inspect, but it does not guarantee that every
macro, surrounding assumption, or semantic dependency has been understood.

\section{Assumptions, conventions, and dependencies}

Assumption manifests and convention registries are the bridge between local
source retrieval and serious mathematical review. An assumption manifest can
declare object kind, shape, symmetry, positive definiteness, invertibility, or
domain-specific requirements near a label or chapter. A convention registry can
record quote conventions, CIP basis signs, measure notation, yield units, or
other choices that change the meaning of a formula.

The dependency graph links labels, assumptions, conventions, packets, and code
references. Its main product use is impact tracing: if a CIP basis sign
convention changes, or if a symbol is reused in two roles, the graph can list
which labels and proof packets should be re-audited. This is diagnostic context
management; it is not an automatic proof repair mechanism.

\section{Packets and claim support}

Proof packets are durable review artifacts built around a label. They collect
source provenance, parser policy, nested audit evidence, high-priority actions,
and a certification boundary. Negative-evidence packets preserve mismatches,
blocked audits, and likely causes so a team can avoid rediscovering the same
failure. Claim-support packets are deliberately separate: they can link a claim
to citations, labels, assumptions, or proof packets, but citation support is not
mathematical proof unless it is tied to a certified scoped obligation.

\section{Matrix IR and code contracts}

The matrix/operator IR is a diagnostic representation for ordered matrix
products and common finance/economics operators: inverse, solve, transpose,
trace, determinant, log determinant, quadratic form, expectation, derivative,
differential, and Jacobian. Its first job is to prevent noncommutative
expressions from being flattened into scalar strings. Unsupported constructs
should remain visible as unresolved IR nodes and should lead to
\texttt{unverified} or \texttt{inconclusive} outcomes.

Code-contract diagnostics complement mathematical audits by checking whether an
implementation exposes shape, dtype, batch-axis, finite-value, or XLA/custom-op
boundaries. These are implementation-quality signals. They can explain why a
state-space, HMC, or neural-solver implementation deserves review, but they do
not establish semantic equivalence to the document.

\chapter{Parser Policy}

MathDevMCP measures parser behavior before relying on parser output for proof
audit. The current lightweight parser is selected for proof audit when it
preserves expected labels and line provenance. LaTeXML is also validated in the
strict LaTeXML profile, but current release routing prefers the provenance and
environment structure available through the current parser.

The policy is deliberately conservative. Parser output becomes dangerous when an
agent treats a nearby equation as if it were the exact source of a claim. The
parser benchmark therefore checks label preservation, environment recognition,
align-like block handling, source spans, include handling, macros, and duplicate
labels. The parser selected for proof-audit routing must preserve the labels the
benchmark expects and expose provenance that a colleague can inspect.

LaTeXML remains valuable because it exercises an independent parser backend and
is closer to a full LaTeX processing stack. In this release it is strict profile
evidence rather than the default proof-audit parser, because the lightweight
parser currently gives the line-oriented provenance and environment structure
used by the downstream audit tools. A future release can change that policy, but
only after benchmark evidence and report examples show that the backend supports
the same workflow.

The owner for this decision is \texttt{parser\_policy.py}, with benchmark
measurements in \texttt{parser\_benchmark.py}. A maintainer should change those
modules before editing this narrative, because the report is downstream of the
policy evidence.

\lstinputlisting[caption={Parser policy summary},basicstyle=\ttfamily\footnotesize]{generated/release_report/parser-policy-summary.txt}

\chapter{Benchmark Gate}

The benchmark gate is the main regression control. It covers consistency,
derivation, workflow, proof-audit, parser-corpus, AST-corpus, typed IR,
industrial review, and release-corpus cases. The gate currently requires every
benchmark case to pass; expected abstentions are counted as quality signals, not
as failures.

The gate is not just a CI badge. It is the release contract that stops the
product from drifting into confident but unsupported behavior. Some benchmark
cases are expected to return \emph{mismatch} or \emph{unverified}; those cases
pass when the tool refuses to overclaim. This is why the report keeps expected
abstentions visible. For a mathematical agent, abstention quality is a product
feature: it tells colleagues that the tool can expose a gap without laundering
the gap into a certificate.

When a new domain is added, the benchmark gate should grow before the marketing
claim grows. The new domain needs at least one parser preservation case, one
operation or code-document consistency case, and one false-confidence or
abstention case if the domain contains assumptions that are easy to miss.

This makes the gate a product-management tool as well as an engineering tool:
it tells the team which claims are backed by repeatable evidence and which
claims still belong in the roadmap.

\lstinputlisting[caption={Benchmark gate summary},basicstyle=\ttfamily\footnotesize]{generated/release_report/benchmark-gate-summary.txt}

\chapter{Specific Domain Tools and How To Use Them}

The domain tools are designed for mathematical finance and economics work where
a single displayed formula hides many smaller obligations. They do not replace
the core proof-audit tools. They help an agent choose a review route, generate
smaller obligations, attach assumptions and conventions, and package the result
for another reviewer.

The current first-wave template catalog is intentionally small:
\path{kalman_loglikelihood_v1}, \path{cip_sdf_sign_v1}, and
\path{hmc_transform_jacobian_v1}. These templates cover three recurring
families in the BayesFilter, DSGE, and CIP monograph workflows: Gaussian
prediction-error likelihoods, CIP/SDF sign-convention reconciliation, and
posterior-geometry transform/Jacobian checks. The roadmap should add more
templates, but the product pattern is already clear.

\section{Template workflow}

A safe domain-template workflow has four steps:

\begin{enumerate}[label=\arabic*.]
  \item locate the candidate label and equation row;
  \item ask for a domain-template suggestion from the label, section path, or
        equation text;
  \item generate template obligations and attach assumptions, conventions, and
        code-contract evidence;
  \item build a proof packet or negative-evidence packet that records what was
        certified, what was diagnostic, and what remains for human review.
\end{enumerate}

\begin{lstlisting}[language=bash,caption={Domain-template discovery and packet workflow}]
PYTHONPATH=src python -m mathdevmcp.cli status-taxonomy

PYTHONPATH=src python -m mathdevmcp.cli domain-templates

PYTHONPATH=src python -m mathdevmcp.cli suggest-domain-templates \
  --label eq:dept-state-space-likelihood \
  --equation-text "logdet innovation covariance solve"

PYTHONPATH=src python -m mathdevmcp.cli generate-template-obligations \
  kalman_loglikelihood_v1 \
  --label eq:dept-state-space-likelihood

PYTHONPATH=src python -m mathdevmcp.cli proof-packet-label \
  eq:dept-state-space-likelihood \
  --root benchmarks/fixtures \
  --summary-only
\end{lstlisting}

The output should be read as a review plan. For example, the Kalman template
generates obligations such as log-det domain, linear-solve residual, quadratic
form shape, and prediction-error decomposition. Each obligation begins as
\texttt{unverified}. It becomes certifying only if a deterministic backend
checks the exact scoped obligation under explicit assumptions.

\section{Domain tool map}

\begin{longtable}{p{0.24\textwidth}p{0.66\textwidth}}
\toprule
Domain family & Current tools and use pattern \\
\midrule
Filtering and state-space likelihoods &
\path{suggest-domain-templates}, \path{generate-template-obligations},
\path{audit_implementation_label}, \path{audit_kalman_recursion},
\path{typed_obligation_label}, \path{proof-packet-label}, and
\path{code_contracts}. Use the Kalman likelihood template to decompose a
prediction-error likelihood
into log-det, solve, quadratic-form, and innovation-covariance obligations.
Then compare the label to code for required operations and use code-contract
diagnostics to check shape, dtype, batch-axis, and finite-value boundaries. \\
CIP, SDF, FX basis, and affine pricing &
\path{cip_sdf_sign_v1}, \path{conventions}, \path{dependency_graph},
\path{claim-support}, and \path{proof-packet-label}. Use the CIP/SDF template
to force quote convention, basis sign convention,
discount-factor definitions, and no-arbitrage identity into explicit review
items. Use the convention registry and dependency graph to identify which
labels must be re-audited after a sign-convention or measure-notation change. \\
DSGE, macro equilibrium, and perturbation &
\path{assumption_manifest}, \path{matrix_ir}, \path{dependency_graph},
\path{typed_obligation_label}, and \path{code_contracts}. Use manifests to
declare steady-state, determinacy, differentiability,
conformability, and shock-process assumptions. Use matrix IR and typed
diagnostics for Kronecker, Sylvester, Jacobian, Hessian, and policy-function
obligations. Future templates should add Euler-equation, Taylor-rule,
Blanchard-Kahn, and perturbation-specific decompositions. \\
HMC, transforms, and posterior geometry &
\path{hmc_transform_jacobian_v1}, \path{matrix_ir},
\path{numeric_diagnostics}, \path{lean-readiness},
\path{proof-packet-label}. Use the HMC transform template to separate
change-of-variables, log-Jacobian,
finite-target, and acceptance-ratio obligations. Use Lean readiness before
asking for formal proof work, and preserve the distinction between approximate
surrogate evidence and exact posterior correction. \\
Particle methods, transport, and neural solvers &
\path{negative-evidence-label}, \path{claim-support},
\path{matrix_ir}, and \path{code_contracts}. Current support is mostly
diagnostic. Use negative-evidence packets to record
failed unbiasedness, resampling, support, or transformed-density checks. Future
templates should cover ESS, differentiable resampling, Sinkhorn/OT objectives,
flow log determinants, HJB residuals, and surrogate-correction boundaries. \\
Literature, assumptions, and empirical claims &
\path{claim-support}, local literature/review fixtures, and proof packets. Use
claim-support packets to classify a statement as an assumption, empirical
regularity, proposed extension, or cited support. Link packets to labels and
proof packets, but do not treat citation support as mathematical proof. \\
\bottomrule
\end{longtable}

\section{Using domain tools in MCP conversations}

The MCP server currently exposes the core conversational checks plus
\path{status_taxonomy} and \path{lean_readiness}. Template and packet commands
are CLI/library-first in this release. A practical agent can still use them in
two ways: call MCP tools for search, lookup, code comparison, typed obligation,
proof audit, release readiness, and Lean readiness; then ask the operator or CI
workflow to run the CLI packet/template commands when a durable artifact is
needed.

\begin{lstlisting}[basicstyle=\ttfamily\scriptsize,caption={MCP-style domain-tool conversation}]
User:
  This CIP chapter changed the sign convention for the basis. What should be
  rechecked?

Agent:
  I will locate the labels and inspect the convention-linked context first.

Tool: search_latex
  query="CIP basis sign convention"

Tool: latex_label_lookup
  label=eq:cip-basis-definition

Agent:
  The next product step is to run the CIP/SDF domain template and convention
  impact report. Those are diagnostic: they should list labels, assumptions,
  and proof packets to re-audit, not claim the revised convention is correct.
\end{lstlisting}

\section{What the domain tools make easier}

For future work like the CIP monograph and DSGE/HMC monograph, the needed
capability list is broader than filtering:

\begin{itemize}
  \item notation and sign-convention management for overloaded finance symbols;
  \item no-arbitrage, SDF, FX forward, basis, and affine-pricing obligations;
  \item Riccati, OU transition, covariance-integral, and yield-from-price
        checks;
  \item Euler-equation, steady-state, perturbation, QZ, Sylvester, and
        Kronecker diagnostics;
  \item Kalman, square-root, SVD, UKF, particle, mixed-frequency, ragged-edge,
        score, Hessian, and finite-safe likelihood checks;
  \item stationarity, ergodicity, Lipschitz, spectral, and filter-stability
        assumption manifests;
  \item posterior geometry, constrained transforms, log-Jacobians, mass
        matrices, HMC/NUTS, and finite-gradient contracts;
  \item transport-map, neural-solver, flow-logdet, HJB-residual, surrogate, and
        delayed-acceptance boundaries;
  \item identification, measurement-system, empirical-design, and observable
        moment packets;
  \item literature and claim-support packets linked to proof packets and local
        review status.
\end{itemize}

The important product rule is the same across all domains: templates make proof
debt visible and repeatable. They do not convert a monograph-scale derivation
into a theorem unless each generated obligation is separately certified.

\part{Colleague Workflows}

\chapter{Workflow 1: Find Mathematical Context}

A colleague begins with a vague request: "Where is the likelihood formula?" or
"Which note defines the Euler equation we used in that experiment?" This is the
moment where an ordinary language model is most tempted to answer from memory.
MathDevMCP's first move is different: search the local LaTeX tree and return
labels, files, line spans, section paths, and text snippets.

\section*{When to use it}

Use \texttt{search-latex} when the colleague has a topic but not a label. It is
especially useful at the beginning of a review, before asking the agent to
compare code or audit a derivation. The search result gives the label that later
commands use as an anchor.

\section*{Command}

The example below searches for the Kalman likelihood and returns the matching
equation label with file and line provenance.

\lstinputlisting[caption={Search workflow output},basicstyle=\ttfamily\scriptsize]{generated/release_report/workflow-search.txt}

\section*{How to read the output}

The important fields are the label, file, line span, score, and text. A high
score is a useful retrieval signal, not a proof that the result is the only
relevant statement. A colleague should inspect the returned section path and
then move to the neighborhood extraction workflow if assumptions or notation are
likely nearby.

\section*{Failure mode}

The main failure mode is semantic overreach. A search hit for "likelihood" may
find the right formula but not the assumptions that make the formula valid. If
the search returns no label, the agent should say so and ask for a narrower
query or a document root, rather than fabricating a source.

\section*{Agent handoff}

An agent handoff should cite the label and provenance from the output rather
than paraphrasing from memory. It should not summarize the equation as verified
until later workflows check implementation terms and derivation evidence.

\chapter{Workflow 2: Read a Labeled Neighborhood}

Once a label is known, the next question is context. Mathematical documents are
rarely self-contained at a single line. A likelihood equation might depend on
notation introduced in the preceding align block. A stability condition might
depend on a boundary convention. A theorem might be harmless in one model class
and false in another.

\section*{When to use it}

Use \texttt{latex-label-lookup} after search and before proof or code
claims. It is the right command when the colleague says, "Show me the surrounding
assumptions" or "What notation does this equation rely on?"

\section*{Command}

The example extracts the neighborhood around the state-space likelihood label.
It gives the selected block plus nearby paragraphs, line spans, and section
metadata.

\lstinputlisting[caption={Labeled neighborhood output},basicstyle=\ttfamily\scriptsize]{generated/release_report/workflow-context.txt}

\section*{How to read the output}

Read the neighborhood as a source packet. The label is the anchor, the file and
line numbers are where a colleague should inspect the original document, and the
paragraphs are the local assumptions or notation the agent may cite. If the
neighborhood omits an assumption that the proof needs, that is a signal to route
the claim to human review.

\section*{Failure mode}

Neighborhoods are local. They do not prove that a global assumption is present
elsewhere in the document. The agent should use them to ground a claim, then say
what is still missing.

\section*{Agent handoff}

A good handoff says: "I found this label and its immediate context; the next
check is whether the implementation carries the required operations." That
handoff sets up the comparison workflow without pretending the context alone is
certification.

\chapter{Workflow 3: Compare Document and Code}

Code--document drift is one of the highest-value release use cases. A colleague
may trust a derivation in a note while the code has silently changed. The
\texttt{audit-implementation-label} workflow asks a smaller, sharper question: do
the required operations from a labeled document block appear in a chosen code
file, and if needed does the implementation expose the requested operation audit
surface?

\section*{When to use it}

Use this workflow when the review question has a concrete code target. It is not
for broad semantic equivalence. It is for surfacing missing terms, suspicious
renamings, or stale implementations that deserve human attention.

\section*{Command}

The example compares the state-space likelihood against a deliberately
incomplete implementation. The document asks for log determinant and inverse or
solve structure; the code has the log determinant but misses the solve term.

\lstinputlisting[caption={Code--document comparison output},basicstyle=\ttfamily\scriptsize]{generated/release_report/workflow-compare.txt}

\section*{How to read the output}

The status is \texttt{mismatch}; that does not mean MathDevMCP proved the code
wrong. It means the configured required term was not found in the inspected code
text. The useful fields are \texttt{matched\_term}, \texttt{missing\_term},
\texttt{code\_path}, and the document provenance. Those fields are enough to
write a precise review ticket.

\section*{Failure mode}

A term check can miss semantic equivalents or be fooled by a misleading
variable name. This is why the report calls it a drift detector, not a proof.
The next step is to inspect the implementation and, if needed, add a benchmark
that captures the intended operation more robustly.

\section*{Agent handoff}

The agent should say: "The document label requires \texttt{solve}; the selected
implementation did not expose that term. Please review whether the quadratic
form is implemented with a numerically appropriate solve or an equivalent
operation." That is useful, bounded, and honest.

\chapter{Workflow 5: Audit a Derivation}

The v2 audit workflow is the product's strongest guard against false confidence.
It combines parser policy, typed obligations, shape diagnostics, numeric
suggestions, backend evidence, and verification-boundary text. Together with
\texttt{typed-obligation-label} and \texttt{check-equality}, it gives an agent a
proof-carrying ladder: extract obligations, inspect typed diagnostics, check
scoped equalities, then summarize only what the evidence supports. It is the
right surface when the colleague asks, "Is this certified, or just plausible?"

\section*{When to use it}

Use \texttt{audit-derivation-v2-label} when the label contains a mathematical
claim whose assumptions matter. Matrix inverses, log determinants, gradients,
expectations, and stability conditions all carry hidden obligations. The audit
exposes those obligations instead of burying them in prose.

\section*{Command}

The example audits the state-space likelihood. It finds high-priority actions:
state or verify invertibility, square-matrix, conformability, log-det domain, and
linear-solve residual constraints.

\lstinputlisting[caption={Derivation audit output},basicstyle=\ttfamily\scriptsize]{generated/release_report/workflow-audit.txt}

\section*{How to read the output}

The key field is the verification boundary. If the status is \texttt{unverified}
or \texttt{inconclusive}, the tool is telling the agent not to claim a proof.
The high-priority actions are the review checklist. A colleague can turn them
into assumptions to state, tests to add, or formal obligations to discharge.
When a reviewer needs the next layer down, \texttt{typed-obligation-label}
localizes dimensional and shape concerns, and \texttt{check-equality} can test a
bounded algebraic step under explicit assumptions.

\section*{Failure mode}

The audit can be verbose. The danger is not verbosity; the danger is summarizing
away the boundary. A release-quality agent must keep the words "unverified",
"diagnostic", or "requires human review" visible when those statuses appear.

\section*{Agent handoff}

The handoff should separate evidence from certification: "The parser located the
label and the audit identified missing matrix constraints. No deterministic
backend certificate is present, so this remains a human-review item."

\chapter{Workflow 6: Construct a Difficult Proof or Derivation}

Many colleague workflows are not just about checking whether code visibly matches
an equation. In a research monograph, the harder task is to build a trustworthy
proof or derivation across a large document set: hundreds of chapters, repeated
notation, hidden assumptions, and intermediate lemmas that may be spread across
several files. This is the setting where a generic agent is most likely to sound
plausible and least likely to be trustworthy.

MathDevMCP is useful here because it gives the colleague a disciplined
proof-carrying workflow rather than a single yes-or-no answer. The agent can
locate candidate labels, pull local neighborhoods, extract typed obligations,
check bounded equalities, route obligations to deterministic or diagnostic
backends, and keep an explicit record of what is proved, what is only supported,
and what remains open.

\section*{When to use it}

Use this workflow when the colleague is trying to construct or extend a difficult
proof, not merely audit one finished equation. Typical examples are matrix
calculus derivations, Kalman/HMC identities, SVD or eigenvalue perturbation
lemmas, multistep Hessian expansions, or chapter-spanning arguments in a large
monograph such as \path{/latex/CIP_monograph/main.tex}. In these settings the
agent should not jump directly to a final formula. It should build a proof ladder
with explicit obligations.

\section*{Why the large-document setting is different}

A 500--700 page monograph behaves differently from a short note.
Definitions are reused across chapters, hypotheses may be stated once and used
much later, and the same symbol may change meaning by section. A colleague needs
an agent that can say:

\begin{enumerate}[label=\arabic*.]
  \item where the current claim is anchored;
  \item which nearby assumptions are visible and which are still missing;
  \item which subclaim has backend evidence;
  \item which step is only heuristic or diagnostic;
  \item what exact prompt the next reviewer should give to continue.
\end{enumerate}

That is the working style recommended by the proof-carrying derivation guide, and
it is the right level of discipline for monograph-scale mathematical work.

\section*{The colleague workflow}

For a difficult proof, the recommended conversation with the agent has seven
turns:

\begin{enumerate}[label=\arabic*.]
  \item identify the target theorem, proposition, or equation label;
  \item retrieve nearby assumptions and notation with provenance;
  \item restate the next proof step as a small proposition;
  \item ask for a proof-audit status before asking for polished prose;
  \item split the argument into typed obligations and bounded equalities;
  \item route each obligation to the most appropriate backend or diagnostic;
  \item ask for a handoff note that lists proved steps, open steps, and the next
        prompt to run.
\end{enumerate}

The core principle is simple: the LLM may organize the derivation, but the proof
boundary must be set by tool evidence. A polished paragraph is not a
certificate.

\section*{Detailed prompt pattern for colleagues}

A strong prompt in the difficult-proof setting should name:

\begin{itemize}
  \item the repository or monograph root;
  \item the label to prove or extend, if known;
  \item the exact subclaim to construct next, not the whole chapter at once;
  \item the assumptions the colleague believes are relevant;
  \item the desired output form: obligations, equality checks, backend evidence,
        or a reviewer handoff note.
\end{itemize}

A weak prompt is: ``Prove the theorem in Chapter 11.''
A much better prompt is:

\begin{quote}
In \path{/latex/CIP_monograph/main.tex}, locate the label for the Gaussian
score differential used in the filtering chapter. Extract nearby assumptions,
restate the next missing step as a small proposition, run a derivation audit,
then tell me which obligations are certified, which are only diagnostic, and
what exact equality or shape checks we should run next.
\end{quote}

That prompt is effective because it forces the agent to retrieve, decompose, and
audit before narrating.

\section*{MCP conversation for proof construction}

\begin{lstlisting}[basicstyle=\ttfamily\scriptsize,caption={MCP conversation for constructing a difficult derivation in a large monograph}]
User:
  In /latex/CIP_monograph/main.tex, help me construct the proof of the
  Gaussian score differential used in the filtering chapter. Do not give me
  polished prose first. I want the proof boundary explicit. Start by finding
  the label, then extract nearby assumptions, then break the next missing step
  into small obligations.

Agent:
  I will first locate the candidate label and inspect its local neighborhood
  before proposing any proof step.

Tool: search_latex
  root=/latex/CIP_monograph
  query="Gaussian score differential filtering likelihood"
  limit=5

Tool result:
  label=eq:cip-gaussian-score-differential
  file=chapters/filtering/gaussian_likelihood.tex
  lines=...

Agent:
  I found a candidate label. I will retrieve the local context and then audit
  the derivation boundary before constructing the next step.

Tool: latex_label_lookup
  root=/latex/CIP_monograph
  label=eq:cip-gaussian-score-differential
  before=1
  after=1

Tool result:
  status=found
  context includes the likelihood equation, solve variable alpha, and local SPD
  assumptions for the covariance matrix

Tool: audit_derivation_v2_label
  root=/latex/CIP_monograph
  label=eq:cip-gaussian-score-differential
  summary_only=true

Tool result:
  status=unverified
  high_priority_actions=[
    "state or verify invertibility",
    "state or verify conformable products",
    "check square trace operand",
    "split the derivation into smaller obligations"
  ]

Agent:
  The current derivation is not certified as one block. I will restate the next
  step as two smaller propositions and localize the typed obligations.

Tool: typed_obligation_label
  root=/latex/CIP_monograph
  label=eq:cip-gaussian-score-differential

Tool result:
  status=consistent
  obligations include solve-shape, trace-shape, and matrix-symmetry concerns

Agent:
  I will now test one bounded algebraic identity instead of the full proof.

Tool: check_equality
  lhs="d alpha"
  rhs="solve(S, dv - dS alpha)"
  assumptions=["S is invertible", "alpha = solve(S, v)", "dS is conformable"]
  backend="sympy"

Tool result:
  status=verified
  evidence=...

Agent:
  We now have one certified substep. The remaining score differential should be
  handled as a separate obligation rather than merged into one paragraph proof.
\end{lstlisting}

\section*{How to work step by step with the agent}

The proof-carrying guide recommends a specific discipline, and colleagues should
expect the agent to follow it.

\subsection*{Step 1: Anchor the claim}

First ask the agent to find the exact label, not just the chapter topic. In a
large monograph, several nearby equations may look similar. The label, file, and
line span are what make the conversation auditable.

\subsection*{Step 2: Retrieve only local context at first}

Ask for the nearby paragraphs before asking for a proof. This prevents the agent
from importing assumptions from memory or from another chapter. The local
context may be incomplete; that incompleteness is itself useful evidence.

\subsection*{Step 3: Restate one small proposition}

Do not ask the agent to prove a page-long Hessian or a full chapter theorem in
one turn. Ask it to restate the next missing step as a small proposition: a solve
differential, a trace identity, a log-determinant differential, a symmetry fact,
or one product-rule expansion.

\subsection*{Step 4: Ask for audit status before prose}

The important intermediate question is not ``Can you explain this?'' but ``Is
this subclaim certified, unverified, inconclusive, or mismatched?'' That wording
causes the agent to preserve the proof boundary.

\subsection*{Step 5: Split obligations aggressively}

If the audit returns \texttt{unverified}, that is usually healthy. It means the
current claim is too large or too implicit. The next request should be to split
the step into smaller obligations. For matrix calculus, that often means:

\begin{enumerate}[label=\alph*.]
  \item a shape or conformability obligation;
  \item a solve or inverse differential identity;
  \item a trace-cyclicity or scalarization step;
  \item a final collection of terms.
\end{enumerate}

\subsection*{Step 6: Route the right obligation to the right backend}

Colleagues should not expect one backend to do everything. A practical routing
pattern is:

\begin{itemize}
  \item \texttt{typed-obligation-label} for dimensions, symmetry, and shape;
  \item \texttt{check-equality} for a bounded symbolic step;
  \item \texttt{audit-derivation-v2-label} for the release-facing summary of the
        whole labeled block;
  \item external finite-difference, AD, Sage, Z3, or Lean checks when the
        proposition is stable enough to justify them.
\end{itemize}

This is exactly the style needed for difficult proofs: the agent should route,
not bluff.

\subsection*{Step 7: End each pass with a handoff note}

At the end of each pass, ask the agent for a concise handoff note with:

\begin{itemize}
  \item the target label;
  \item the proposition checked this pass;
  \item assumptions used explicitly;
  \item statuses of each tool call;
  \item certified substeps;
  \item open obligations;
  \item the exact next prompt to run.
\end{itemize}

That handoff note is what lets another colleague continue after a reset without
reconstructing the same uncertainty from scratch.

\section*{What a good agent answer should look like}

For a difficult proof, a good answer is not a long elegant derivation on the
first turn. A good answer says something like:

\begin{quote}
I found label \path{eq:cip-gaussian-score-differential} in the filtering
chapter. The local neighborhood supports the solve notation and SPD assumption,
but the full derivation remains \path{unverified} as one block. One substep,
\path{d alpha = solve(S, dv - dS alpha)}, is verified under the explicit
invertibility and conformability assumptions. The next step is to isolate the
trace-form differential and check it separately.
\end{quote}

That answer is useful because it distinguishes retrieval, certification, and the
next proof obligation.

\section*{Failure modes in monograph-scale proof construction}

The most common failure modes are:

\begin{enumerate}[label=\arabic*.]
  \item asking for a whole difficult proof in one turn;
  \item accepting nearby notation as if it were a proof;
  \item failing to restate assumptions explicitly;
  \item merging shape issues and algebraic issues into one giant claim;
  \item treating \texttt{unverified} as permission to write polished certainty;
  \item forgetting that a symbol may shift meaning across chapters;
  \item treating a symbolic toy check as a theorem for all dimensions.
\end{enumerate}

Colleagues should treat these as workflow bugs, not only mathematical bugs.

\section*{Boundary}

MathDevMCP helps construct difficult proofs by keeping the work auditable. It
does not replace expert judgment, and it does not turn a monograph-scale proof
into one automatic backend call. Its value is that it lets a colleague and an
agent collaborate in small certifiable steps, with explicit provenance and an
explicit record of what remains open.

\chapter{Workflow 7: Build an Implementation Brief}

The implementation brief is the handoff format for a busy colleague. It bundles
search, label selection, context extraction, consistency checking, and optional
derivation evidence into one object. It is the report a teammate can paste into
a review thread before opening the code.

\section*{When to use it}

Use \texttt{implementation-brief} when another person or agent needs a grounded
starting point. It is especially useful before a pull request review, a
maintenance ticket, or a focused refactor of mathematical code.

\section*{Command}

The example asks for a state-space likelihood brief against the same incomplete
implementation used above. The brief selects the relevant label, includes
context, and reports the consistency mismatch.

\lstinputlisting[caption={Implementation brief output},basicstyle=\ttfamily\scriptsize]{generated/release_report/workflow-brief.txt}

\section*{How to read the output}

The brief's \texttt{selected\_label} tells the reviewer what document claim was
used. The \texttt{checks.consistency.status} field tells whether the required
terms appeared. The context excerpt gives the exact mathematical block. This is
enough for a reviewer to reproduce the check without trusting the agent's
memory.

\section*{Failure mode}

The brief is only as good as the query and required terms. If the query selects
the wrong label, or if the required terms are too weak, the brief can be
uninformative. The right response is to rerun with a sharper query or a better
term set, not to inflate the conclusion.

\section*{Agent handoff}

A good handoff reads like a compact review note: selected label, code file,
status, missing terms, and next action. It should avoid broad claims such as
"the implementation is wrong" unless a human review or stronger backend evidence
supports that claim.

\part{Case Studies}

\chapter{Workflow 8: Kalman State-Space Likelihood}

\section*{Colleague scenario}

A colleague is reviewing a Kalman filter implementation before using it in a
forecasting experiment. The note defines a Gaussian likelihood with
\(\log\det S_t\) and \(S_t^{-1}\) terms. The implementation has been optimized
with a scan loop, and the reviewer wants to know whether the code still exposes
the solve-like operation implied by the derivation.

\section*{Fixture and command}

The public fixture uses the state-space document and a deliberately incomplete
missing-solve implementation. The command compares the likelihood label against
required terms \path{logdet,solve}.

\section*{MCP conversation}

\begin{lstlisting}[basicstyle=\ttfamily\scriptsize]
User:
  Check whether the state-space likelihood label matches the Kalman code.
  Look specifically for logdet and solve-like structure.
Agent:
  I will anchor the review on the document label before judging the code.
Tool: search_latex(root, "state space likelihood logdet solve")
Tool: audit_implementation_label(root, "eq:dept-state-space-likelihood",
                         code, required_terms=["logdet", "solve"],
                         before=1, after=1)
Agent:
  The result is mismatch: logdet is visible, solve is missing. This is a
  drift signal, not a proof of invalidity. Review whether an equivalent
  factorization implements the quadratic solve.
\end{lstlisting}

\lstinputlisting[caption={Kalman likelihood comparison evidence},basicstyle=\ttfamily\scriptsize]{generated/release_report/case-kalman-compare.txt}

\section*{Output to inspect}

The output is intentionally uncomfortable: status \texttt{mismatch}, matched
term \texttt{logdet}, and missing term \texttt{solve}. That is exactly the kind of result that
makes the product useful. It does not bury a missing numerical operation under a
friendly summary. It gives a reviewer the label, file, line span, code path, and
missing term.

\section*{Interpretation}

The mathematical issue is numerical, not merely textual. A likelihood written
with \(S_t^{-1}\) should usually be implemented through a solve or factorization,
not by casually dropping the quadratic solve. The tool cannot prove the code is
wrong from the missing token alone, but it can make the review precise: inspect
the innovation quadratic form and decide whether a stable solve is present under
another name.

\section*{Next action}

The colleague should open the implementation, check how the quadratic form is
computed, and add or fix a test that exercises conditioning of \(S_t\). If the
code intentionally uses an equivalent factorization, the benchmark should be
updated with an operation extractor or required term that recognizes the chosen
idiom.

\section*{Boundary}

MathDevMCP reports a drift signal, not a theorem. It does not certify the
Kalman likelihood, and it does not prove the implementation is invalid. It
prevents an agent from saying "the code matches the document" when a required
solve-like operation is not visible.

\chapter{HMC Leapfrog and Hamiltonian Flow}

\section*{Colleague scenario}

A Bayesian computation colleague is checking an HMC kernel. The document gives a
leapfrog update and a Hamiltonian. The code has gradients, half-steps, kinetic
energy, and a returned Hamiltonian value. The review question is not "does this
sampler converge?" It is narrower: does the implementation visibly carry the
documented leapfrog and Hamiltonian structure?

\section*{Fixture and command}

The fixture uses \path{doc_department_bayesian_hmc.tex} and
\path{doc_department_hmc_jax.py}. The comparison asks for \path{grad}
and \path{hamiltonian} under label \path{eq:dept-hmc-leapfrog}.

\section*{MCP conversation}

\begin{lstlisting}[basicstyle=\ttfamily\scriptsize]
User:
  Does the HMC implementation visibly carry the leapfrog and Hamiltonian
  structure from the note?
Agent:
  I will compare the leapfrog label against the HMC code for gradient and
  Hamiltonian terms.
Tool: audit_implementation_label(root, "eq:dept-hmc-leapfrog", code,
                         required_terms=["grad", "hamiltonian"])
Agent:
  The structural check is consistent. I can say the selected code is not
  obviously stale with respect to those terms, but I cannot certify detailed
  balance, acceptance behavior, or convergence from this check.
\end{lstlisting}

\lstinputlisting[caption={HMC leapfrog comparison evidence},basicstyle=\ttfamily\scriptsize]{generated/release_report/case-hmc-compare.txt}

\section*{Output to inspect}

The status is \texttt{consistent}: both required terms appear. The output also keeps
the provenance of the align block so a colleague can see the three leapfrog
lines rather than trusting a paraphrase.

\section*{Interpretation}

This is a positive structural check. It says the obvious ingredients for the
documented leapfrog are present in the inspected code. It is useful before a
deeper review because it quickly rules out a stale file that lacks gradients or
Hamiltonian energy entirely.

\section*{Next action}

The next review step is to inspect signs, step ordering, mass matrix semantics,
and Metropolis correction. If the claim is about stationarity or convergence,
route it to human review or a dedicated proof/audit workflow; do not treat a
term match as a Markov-chain proof.

For a release note or pull-request handoff, the useful sentence is: the code is
not obviously stale with respect to the documented leapfrog structure, but the
probabilistic correctness argument remains outside this check.

\section*{Boundary}

MathDevMCP does not certify HMC correctness, detailed balance, ergodicity, or
acceptance-rate behavior here. It confirms that specific document-grounded
operations are visible and keeps the remaining probabilistic claims outside the
certificate boundary.

Maintainers adding richer HMC coverage should add a seeded sign-error or missing
Metropolis-correction case before changing the release claim.

\chapter{Macro Filter Multi-File Corpus}

\section*{Colleague scenario}

The macro-filter case resembles a real internal note: a main LaTeX file defines
macros, includes a model file, and repeats notation across state equations. A
colleague has an implementation that computes prediction and likelihood terms
but may have omitted the Kalman gain update. The question is whether the tool
can follow the included document and expose the omission.

\section*{Fixture and command}

The fixture uses a macro-filter main document, an included model file, and a
deliberately incomplete missing-gain implementation. The comparison below
checks the likelihood label for \path{logdet}, \path{solve}, and \path{K_t}.

\section*{MCP conversation}

\begin{lstlisting}[basicstyle=\ttfamily\scriptsize]
User:
  This macro filter is split across included TeX files. Check whether the
  implementation has the likelihood terms and the Kalman gain update.
Agent:
  I will let the parser resolve the included model file, then compare the
  selected label to the implementation.
Tool: search_latex(root, "macro filter likelihood Kalman gain")
Tool: audit_implementation_label(root, "eq:macro-filter-likelihood", code,
                         required_terms=["logdet", "solve", "K_t"])
Agent:
  The comparison finds logdet and solve but not K_t. The likely review
  question is whether this function is intentionally prediction-only or has
  dropped the gain/update step.
\end{lstlisting}

\lstinputlisting[caption={Macro filter comparison evidence},basicstyle=\ttfamily\scriptsize]{generated/release_report/case-macro-filter-compare.txt}

\section*{Output to inspect}

The output reports \texttt{mismatch}: \texttt{logdet} and \texttt{solve} are
present, while \texttt{K\_t} is missing. The provenance points into the
included model file, not only the main file, which matters because many real
documents are modular.

\section*{Interpretation}

This case demonstrates two product behaviors at once. First, the parser handles
the multi-file LaTeX structure well enough to locate the model equation. Second,
the code comparison can distinguish a likelihood calculation that contains
linear algebra terms from a complete filter update that carries the gain.

That distinction is exactly where colleagues lose time in practice: a partial
routine can look mathematically serious because it contains the log determinant
and solve, while still omitting the update needed by the model description.

\section*{Next action}

The colleague should inspect whether the implementation intentionally returns a
prediction-only likelihood or whether it should update state and covariance. If
the implementation is intentionally partial, the document or function name
should say so. If not, add the missing gain/update logic and a regression test.

If this pattern recurs in a project, make \texttt{K\_t} or the chosen
gain-update idiom part of the domain benchmark rather than relying on a one-off
review comment.

\section*{Boundary}

The tool does not infer all macro semantics or prove the filter recursion. It
surfaces a targeted mismatch in a realistic multi-file setting and leaves the
modeling decision to the reviewer.

\chapter{DSGE Euler Equation}

\section*{Colleague scenario}

A macro-finance colleague has a compact Euler equation in the notes and a small
residual function in code. The review question is whether the implementation
still computes an expectation of a stochastic discount factor times a return,
not whether the model is calibrated, solved, or empirically valid.

\section*{Fixture and command}

The fixture uses \path{doc_sanitized_dsge_macro_finance.tex} and
\path{doc_sanitized_dsge_macro_finance.py}. The comparison checks
\path{eq:dept-euler-equation} for \path{expectation} and
\path{euler_residual}.

\section*{MCP conversation}

\begin{lstlisting}[basicstyle=\ttfamily\scriptsize]
User:
  In the DSGE example, check that the code still represents the Euler
  residual from the note. Do not make calibration claims.
Agent:
  I will compare the Euler-equation label to the residual code for the
  configured structural terms.
Tool: audit_implementation_label(root, "eq:dept-euler-equation", code,
                         required_terms=["expectation", "euler_residual"])
Agent:
  The comparison is consistent for those terms. The code carries the named
  residual structure, but calibration, timing, and equilibrium validity remain
  outside this evidence.
\end{lstlisting}

\lstinputlisting[caption={DSGE Euler comparison evidence},basicstyle=\ttfamily\scriptsize]{generated/release_report/case-dsge-compare.txt}

\section*{Output to inspect}

The status is \texttt{consistent}, and the matched terms show that the implementation
contains both the expectation operator and the residual concept. The provenance
points to the Euler equation in the sanitized macro-finance fixture.

\section*{Interpretation}

This is the kind of small but valuable check that prevents stale glue code. It
does not decide the economic model; it ensures that the code being reviewed at
least carries the structural ingredients named in the document.

The value is speed with restraint. The reviewer can quickly move past the
"wrong file" and "missing residual" class of errors while keeping the economic
interpretation under expert review.

\section*{Next action}

The colleague should next review calibration assumptions, timing conventions,
state definitions, and units. If those are part of the release claim, add
fixture labels and expected abstentions that force the tool to report missing
assumptions.

For production use, this is where a team should add project-specific labels for
household timing, asset returns, and shock normalization if those conventions
are common sources of errors.

\section*{Boundary}

MathDevMCP does not solve the DSGE model, estimate parameters, or validate an
equilibrium. It gives document-grounded evidence that the inspected residual
function has the expected structural terms.

\chapter{Stochastic Volatility Likelihood}

\section*{Colleague scenario}

A colleague is reviewing a stochastic-volatility likelihood implementation. The
document has a transition equation for latent volatility and a conditional
likelihood for observations. The code computes an innovation-like likelihood
term, and the reviewer wants to check that the expected likelihood structure is
visible.

\section*{Fixture and command}

The fixture uses \path{doc_sanitized_stochastic_volatility.tex} and
\path{doc_sanitized_stochastic_volatility.py}. The comparison checks
\path{eq:dept-sv-likelihood} for \path{innovation} and
\path{posterior_or_likelihood}.

\section*{MCP conversation}

\begin{lstlisting}[basicstyle=\ttfamily\scriptsize]
User:
  Check whether this stochastic-volatility likelihood code has the likelihood
  path described in the document.
Agent:
  I will compare the likelihood label against the code for innovation and
  likelihood/posterior structure.
Tool: audit_implementation_label(root, "eq:dept-sv-likelihood", code,
                         required_terms=["innovation",
                                         "posterior_or_likelihood"])
Agent:
  The selected structure is visible. I would still route latent-prior,
  Jacobian, and inference-method assumptions to human review.
\end{lstlisting}

\lstinputlisting[caption={Stochastic volatility comparison evidence},basicstyle=\ttfamily\scriptsize]{generated/release_report/case-stochastic-volatility-compare.txt}

\section*{Output to inspect}

The status is \texttt{consistent}; both required terms are found. The output gives the
document label, code file, matched terms, and the likelihood equation excerpt.

\section*{Interpretation}

This result is useful because it confirms the code has a likelihood-bearing path
that corresponds to the document. It also tells the reviewer what remains out of
scope: latent prior assumptions, filtering/inference strategy, Jacobians, and
parameter constraints are not certified by this term check.

The example is intentionally narrow. It is strong enough to catch a stale file
with no likelihood path, but too weak to approve a full inference algorithm.

\section*{Next action}

The colleague should inspect whether the transition variance, parameter bounds,
and inference algorithm match the intended model. If missing-Jacobian or prior
review is a recurring issue, add a seeded false-confidence case to the benchmark
corpus.

Teams using particle filters, variational approximations, or MCMC for this
model should add separate fixtures for those inference choices, because the
likelihood term alone will not capture them.

\section*{Boundary}

MathDevMCP does not certify the stochastic-volatility likelihood or posterior
inference. It confirms selected structure and keeps latent-model assumptions in
the human-review lane unless more evidence is supplied.

\chapter{SDE and PDE Numerics}

\section*{Colleague scenario}

A numerical methods colleague is checking whether a simulation step reflects the
Euler--Maruyama update in the note and whether the code carries an explicit
stability condition. The aim is to catch silent drift between mathematical
discretization and implementation before experiments are trusted.

\section*{Fixture and command}

The fixture uses \path{doc_sanitized_sde_pde_numerics.tex} and
\path{doc_sanitized_sde_pde_numerics.py}. The comparison checks
\path{eq:dept-euler-maruyama} for \path{drift}, \path{diffusion}, and
\path{stability_condition}.

\section*{MCP conversation}

\begin{lstlisting}[basicstyle=\ttfamily\scriptsize]
User:
  Before we trust these simulations, check whether the code carries the
  Euler-Maruyama drift, diffusion, and stability hook from the note.
Agent:
  I will anchor on the Euler-Maruyama label and inspect the selected code.
Tool: audit_implementation_label(root, "eq:dept-euler-maruyama", code,
                         required_terms=["drift", "diffusion",
                                         "stability_condition"])
Agent:
  The structural terms are present. This supports an implementation-screening
  pass, not a convergence or stability proof.
\end{lstlisting}

\lstinputlisting[caption={SDE/PDE numerics comparison evidence},basicstyle=\ttfamily\scriptsize]{generated/release_report/case-sde-pde-compare.txt}

\section*{Output to inspect}

The status is \texttt{consistent}: the implementation exposes drift, diffusion, and a
stability-condition term. The provenance points to the Euler--Maruyama formula.

\section*{Interpretation}

This is a structural release check, not a numerical analysis proof. It tells the
reviewer that the code carries the named pieces of the discretization. It does
not tell the reviewer that the step size is appropriate for every SDE, that the
PDE residual is stable, or that convergence assumptions hold.

That boundary makes the check useful in a lab setting: it catches missing drift,
diffusion, or stability hooks before anyone spends time interpreting simulated
paths as evidence about the continuous model.

\section*{Next action}

The colleague should inspect the time-step bound, units, boundary conditions,
and randomness convention. If a stability argument is important to the release,
add a proof-audit or typed-obligation example that forces missing assumptions to
surface.

For code review, the best follow-up is often a small stress test near the
documented stability limit, because such tests expose convention mismatches
that token checks cannot see.

\section*{Boundary}

MathDevMCP does not prove convergence or stability. It keeps the implementation
review anchored to the documented discretization and makes missing terms visible
early.

\chapter{ML and LLM Objective Functions}

\section*{Colleague scenario}

An ML colleague is reviewing objective code after several training-loop edits.
The document states a negative log-likelihood loss and its gradient. The code
uses an automatic differentiation function named \texttt{grad}, but the report
is asked to check for the more explicit term \texttt{gradient}. This is a useful
stress test for term-level review.

\section*{Fixture and command}

The fixture uses \path{doc_sanitized_ml_llm_objective.tex} and
\path{doc_sanitized_ml_llm_objective.py}. The comparison checks
\path{eq:dept-ml-gradient} for \path{gradient} and \path{loss}.

\section*{MCP conversation}

\begin{lstlisting}[basicstyle=\ttfamily\scriptsize]
User:
  The ML code uses autodiff. Check whether the documented gradient objective
  is visible, but do not silently equate grad with the full derivation.
Agent:
  I will run the configured term check and report literal matches.
Tool: audit_implementation_label(root, "eq:dept-ml-gradient", code,
                         required_terms=["gradient", "loss"])
Agent:
  The result is mismatch: loss is visible, gradient is not matched exactly.
  The code may use grad or another idiom, so the next step is to decide
  whether the benchmark should recognize that operation explicitly.
\end{lstlisting}

\lstinputlisting[caption={ML objective comparison evidence},basicstyle=\ttfamily\scriptsize]{generated/release_report/case-ml-objective-compare.txt}

\section*{Output to inspect}

The status is \texttt{mismatch}: \texttt{loss} is present, but
\texttt{gradient} is missing. The code does contain \texttt{grad} and
\texttt{gradient\_value}, yet the configured required term is not present
exactly. That is not a failure of honesty; it is a reminder that term checks are
intentionally literal.

\section*{Interpretation}

This example is good for selling the product because it shows both value and
limits. The tool catches that the configured review vocabulary does not line up
with the code. A reviewer can decide whether to accept \texttt{grad} as the
implementation idiom, change the required term list, or add operation extraction
that recognizes automatic differentiation.

\section*{Next action}

The colleague should review sign conventions, reduction over the batch, and the
automatic-differentiation call. If the code is correct, update the benchmark to
look for the intended operation rather than a single English token.

\section*{Boundary}

MathDevMCP does not prove the training objective is correct, and it does not
know that \texttt{grad} is semantically equivalent to the full documented
gradient in all contexts. It reports the literal evidence and asks the reviewer
to sharpen the contract.

\chapter{Bayesian ELBO and Variational Inference}

\section*{Colleague scenario}

A Bayesian ML colleague is reviewing a variational-inference implementation. The
document names the ELBO and a reparameterization gradient. The code computes an
\texttt{elbo\_value}, but the configured term check asks for \texttt{elbo}
exactly. This is
another realistic review problem: naming conventions are close but not
identical.

\section*{Fixture and command}

The fixture uses \path{doc_sanitized_bayesian_elbo_vi.tex} and
\path{doc_sanitized_bayesian_elbo_vi.py}. The comparison checks
\path{eq:dept-elbo} for \path{expectation} and \path{elbo}.

\section*{MCP conversation}

\begin{lstlisting}[basicstyle=\ttfamily\scriptsize]
User:
  Check the ELBO implementation. If the code only has a nearby name such as
  elbo_value, keep that distinction visible.
Agent:
  I will run the literal configured review terms against the ELBO label.
Tool: audit_implementation_label(root, "eq:dept-elbo", code,
                         required_terms=["expectation", "elbo"])
Agent:
  The result is mismatch: expectation is present, elbo is not matched exactly.
  I see clues that may be relevant, but I will not turn those clues into a
  certification without a sharper operation extractor or human review.
\end{lstlisting}

\lstinputlisting[caption={ELBO comparison evidence},basicstyle=\ttfamily\scriptsize]{generated/release_report/case-elbo-compare.txt}

\section*{Output to inspect}

The status is \texttt{mismatch}: \texttt{expectation} is present while
\texttt{elbo} is missing. The extra terms include \texttt{elbo\_value} and
\texttt{reparameterization\_gradient}, which gives the reviewer a useful clue
without silently converting that clue into a match.

\section*{Interpretation}

This is a disciplined false-confidence example. A loose agent might say "ELBO is
there" because it sees \texttt{elbo\_value}. MathDevMCP reports the exact configured
term result and leaves the interpretation visible. That lets the team decide
whether token matching is enough or whether AST/operation extraction should be
extended for this domain.

The point is not to embarrass the checker for being literal. The point is to
make the review contract visible enough that the team can improve it deliberately
instead of accepting a fuzzy match.

\section*{Next action}

The colleague should inspect entropy terms, support assumptions, and whether
the reparameterization path matches the document. If the naming convention is
accepted, tune the release evidence so the benchmark captures the intended
semantic operation.

For future work, the natural extension is an operation extractor that recognizes
ELBO assembly patterns rather than only the spelling of one variable.

\section*{Boundary}

MathDevMCP does not certify the ELBO derivation, variational-family support, or
gradient estimator validity. It gives a reproducible review signal and keeps
the proof boundary explicit.

\chapter{Computational Physics MCMC}

\section*{Colleague scenario}

A computational physics colleague is checking MCMC code that computes a
Hamiltonian acceptance ratio. The document states an acceptance formula, and the
code computes Hamiltonian energies and a log acceptance probability. The review
question is whether those core ingredients are visible.

\section*{Fixture and command}

The fixture uses the sanitized computational-physics MCMC document and its
paired code file. The comparison checks the acceptance-ratio label for
\path{hamiltonian} and \path{gradient}.

\section*{MCP conversation}

\begin{lstlisting}[basicstyle=\ttfamily\scriptsize]
User:
  Does the MCMC implementation carry the Hamiltonian and gradient ingredients
  from the acceptance-ratio note?
Agent:
  I will compare the acceptance-ratio label to the code for those structural
  terms only.
Tool: audit_implementation_label(root, "eq:dept-acceptance-ratio", code,
                         required_terms=["hamiltonian", "gradient"])
Agent:
  The selected terms are present. This is useful screening evidence, but it
  does not certify reversibility, detailed balance, or ergodicity.
\end{lstlisting}

\lstinputlisting[caption={Computational physics MCMC comparison evidence},basicstyle=\ttfamily\scriptsize]{generated/release_report/case-physics-mcmc-compare.txt}

\section*{Output to inspect}

The status is \texttt{consistent}: Hamiltonian and gradient terms are found. The output
also shows the acceptance-ratio label and document excerpt.

\section*{Interpretation}

This is a positive release-facing check. It confirms the implementation has the
Hamiltonian energy and gradient structure expected by the note. It does not
settle detailed balance, reversibility, or convergence. Those require stronger
mathematical arguments and domain review.

This is the right level of automation for an implementation-screening pass:
reject obviously stale code early, then spend human time on the properties that
actually require statistical mechanics or MCMC expertise.

\section*{Next action}

The colleague should inspect proposal symmetry, momentum refresh, Metropolis
correction, and numerical stability. If those are part of the release claim,
they need explicit fixtures and expected abstention cases.

If the team later wants the tool to discuss reversibility, add a dedicated
derivation-audit example with an explicit verification boundary rather than
stretching this term check beyond its contract.

\section*{Boundary}

MathDevMCP does not certify an MCMC sampler. It shows that the code and document
share selected structural ingredients and then stops before overclaiming
ergodicity or correctness.

\chapter{Private Corpus Validation}

\section*{Colleague scenario}

The public fixtures are useful, but an industrial release also needs to prove
that the machinery can validate private or sanitized department material without
committing it. The private-corpus workflow is designed for that trust boundary.
The manifest may contain sensitive local paths, while the report must remain
safe to share.

\section*{Fixture and command}

For this autonomous release, the evidence uses an external sanitized corpus
created outside the repository. It includes six private-sanitized entries across
DSGE, stochastic volatility, SDE/PDE numerics, ML objectives, ELBO/VI, and
computational physics MCMC. A real deployment replaces that manifest with an
approved private manifest using the same environment variable and validator.

\section*{MCP conversation}

\begin{lstlisting}[basicstyle=\ttfamily\scriptsize]
User:
  Validate the private corpus manifest, but do not reveal private paths in the
  answer.
Agent:
  I will validate the manifest and report only redacted release evidence.
Tool: validate_release_corpus(root, private_manifest="/secure/.../corpus.json")
Tool: release_readiness(root, profile="full")
Agent:
  The private manifest is loaded and release-gated entries validate. The
  normal report redacts private roots. If a path inside the git checkout were
  present, I would report it as a release finding rather than proceeding.
\end{lstlisting}

\lstinputlisting[caption={Private corpus validation summary},basicstyle=\ttfamily\footnotesize]{generated/release_report/private-corpus-summary.txt}

\section*{Output to inspect}

The summary reports \texttt{consistent}, zero findings, redacted private paths,
and parser reports selected for proof audit. The important part is not the
temporary directory. The important part is that the validator used unredacted
local paths internally while emitting redacted report output.

\section*{Interpretation}

This closes the release blocker for private-corpus machinery without pretending
that real confidential department documents were committed or inspected by this
repository. It demonstrates that an operator can supply an external manifest and
get release-gated parser evidence with safe output.

\section*{Next action}

Before a department deployment, an operator should replace the sanitized corpus
with an approved private manifest, verify coverage by domain, run
\path{scripts/validate_private_corpus.sh}, and record whether the real
private corpus covers the same domains as the sanitized release corpus.

That record matters for sales and governance. It lets a manager say, "The
release machinery has been tested here; our department-specific claim begins
only after we point it at our approved private material."

\section*{Boundary}

This report's private evidence is sanitized external evidence. It does not claim
that unredacted department-private documents were reviewed. The privacy rule is
simple: populated private manifests stay outside git, normal reports redact
paths, and any private path inside the checkout is a release finding.

The code owner for this boundary is \texttt{release\_corpus.py}; the release
profile owner is \texttt{release\_policy.py}. Both should be reviewed before
changing the private-corpus claim.

\part{Security, Operations, and Maintenance}

\chapter{Security and Privacy}

Security in MathDevMCP starts from a local-first assumption: mathematical
documents, private source trees, and code roots remain on the user's machine.
The product does not need to upload documents to a service in order to search
LaTeX labels, compare code terms, or build a release corpus report. This is a
major part of the industrial value proposition for colleagues who work with
unpublished notes, grant material, confidential model code, or department
calibration files.

\section{Private corpus boundary}

Private documents must stay outside the repository. A populated private
manifest may contain sensitive paths and should remain external. Normal
MathDevMCP reports redact private document roots and code roots through
\texttt{release\_corpus.py}. Governance validation rejects private paths inside
the checkout. Evidence bundles for review should be stored outside git unless
they are deliberately reduced to safe report snippets.

The privacy contract has three levels. Public fixtures are safe to commit and
are used for repeatable tests. External sanitized fixtures may be used to prove
release machinery without revealing real documents. Real private manifests are
operator-supplied and should never appear in git. The report uses the second
level and says so plainly.

\section{Command and backend governance}

External commands are bounded by governance policy. \texttt{governance.py}
checks that MathDevMCP subprocess calls include explicit timeouts, and the
policy allowlist names the external tools the project expects: LaTeXML, Pandoc,
Lean, Lake, and Sage. Backend failures are reported as structured evidence
rather than raw shell surprises.

This matters for agent use. An MCP tool call should not become a hidden shell
escape hatch. It should return a typed report with a status, findings, and
provenance. When a backend is unavailable, times out, or returns diagnostic-only
evidence, the result should make that boundary visible.

\section{Review checklist}

Before sharing a release bundle, scan generated report evidence for absolute
private paths, manifest filenames, and local checkout paths. Run
\path{scripts/audit_release_report_substance.sh} and the release-readiness
profile with the intended manifest. If a scan finds an unredacted private path,
treat it as a release blocker, not a cosmetic documentation issue.

\chapter{Operations}

Operations are intentionally command driven. A colleague should be able to
clone the repository, install the package, run doctor checks, and obtain a
release-readiness report without opening a notebook or editing the source. The
normal operator sequence is:

\begin{lstlisting}[language=bash]
PYTHONPATH=src python -m mathdevmcp.cli doctor
scripts/backend_env_doctor.sh "$PWD"
MATHDEVMCP_REQUIRE_LATEXML=1 scripts/validate_latexml_backend.sh "$PWD"
PYTHONPATH=src python -m mathdevmcp.cli benchmark-gate --root "$PWD"
PYTHONPATH=src python -m mathdevmcp.cli release-readiness --root "$PWD" --profile base
\end{lstlisting}

For full release validation, add:

\begin{lstlisting}[language=bash]
export MATHDEVMCP_PRIVATE_CORPUS_MANIFEST=/secure/local/path/corpus.json
scripts/validate_private_corpus.sh "$PWD"
PYTHONPATH=src python -m mathdevmcp.cli release-readiness --root "$PWD" --profile full
\end{lstlisting}

\section{Interpreting a run}

Read the commands in layers. \texttt{doctor} describes local capability:
Python, LaTeXML, Pandoc, Lean, Sage, SymPy, and LeanDojo. Backend env doctor
checks the isolated environment used for LeanDojo. LaTeXML validation is strict
only when the release profile requires it. The benchmark gate checks product
behavior; release readiness combines those results with governance, parser
policy, private-corpus status, and git state.

A dirty working tree is a caveat because release reports should be tied to a
commit. It is normal during documentation editing, but it must be cleared or
accepted explicitly before a tag. Missing private corpus evidence is a caveat
for the base profile and a blocker for the private-corpus and full profiles.

\section{Environment layout}

Lean may be available through \texttt{elan} in the user environment. LeanDojo is
kept in \texttt{mathdevmcp-backends} and invoked through
\texttt{MATHDEVMCP\_BACKEND\_CONDA\_ENV}. This split is deliberate: it prevents
the backend dependency stack from destabilizing the main environment while
still making backend evidence available for strict profiles.

\section{Release cadence}

For a release branch, regenerate evidence after any change to parser policy,
release corpus metadata, benchmark cases, release policy, or user-facing
workflow examples. Rebuild the PDF after evidence regeneration. Run the
substance audit before final review so page count cannot hide thin chapters.

\chapter{Maintainer Guide}

Maintain release behavior by changing library code first, then CLI/MCP wrappers,
then docs and evidence. When adding a new benchmark category, update the case
builder, runner, summary tests, release report evidence, and release policy if
the category becomes mandatory. When adding a new private corpus entry, keep the
entry outside git and check redaction before sharing output.

The maintainer rule of thumb is simple: a release claim should have one owning
module, one CLI or MCP surface, at least one test, and one report example if it
is colleague-facing. If any of those four pieces is missing, the claim is not
ready to sell as an industrial feature.

The highest-risk modules are:

\begin{itemize}
  \item \texttt{benchmarks.py}, for benchmark totals and gate behavior;
  \item \texttt{release\_policy.py}, for profile blockers and caveats;
  \item \texttt{release\_corpus.py}, for privacy redaction and manifest shape;
  \item \texttt{parser\_policy.py}, for proof-audit parser routing;
  \item \texttt{proof\_audit\_v2.py}, for verification-boundary reporting;
  \item \texttt{status\_taxonomy.py}, for public status and substatus
        vocabulary;
  \item \texttt{equation\_locator.py} and \texttt{latex\_index.py}, for
        source-local equation provenance;
  \item \texttt{assumption\_manifest.py}, \texttt{conventions.py}, and
        \texttt{dependency\_graph.py}, for diagnostic context management;
  \item \texttt{matrix\_ir.py}, for noncommutative operator diagnostics;
  \item \texttt{proof\_packet.py}, \texttt{negative\_evidence.py}, and
        \texttt{claim\_support.py}, for durable review artifacts;
  \item \texttt{domain\_templates.py}, for governed domain decompositions;
  \item \texttt{code\_contracts.py}, for implementation-boundary diagnostics;
  \item \texttt{lean\_readiness.py}, for direct Lean, Lake, and LeanDojo
        readiness separation;
  \item \texttt{mcp\_facade.py} and \texttt{cli.py}, for public tool surfaces.
\end{itemize}

\section{Change workflow}

For a behavior change, start with the library function and its direct tests.
Then update the CLI command if arguments or JSON shape changed. Then update the
MCP facade so agents receive the same behavior. Only then regenerate report
evidence and edit narrative documentation. This order keeps prose downstream of
executable behavior.

\section{Contract hygiene}

Do not rename statuses casually. Words such as \texttt{mismatch},
\texttt{unverified}, \texttt{inconclusive}, and \texttt{ready\_with\_caveats}
are part of the user contract. A rename can break tests, release snippets,
agent prompts, and colleague training material at the same time.

\section{Documentation hygiene}

Documentation changes that touch release claims should run the same gates as
small code changes: generated evidence, report-substance audit, PDF build, path
leak scan, and release-readiness check. The report is a sales document, but it
is also a reproducibility artifact.

\chapter{Limitations and Accepted Boundaries}

The product is conservative by design. It is useful because it knows when to
stop. It does not convert every mathematical statement to a theorem prover. It
does not guarantee semantic equivalence between arbitrary prose and code. It
does not make private data safe to commit. It does not make numeric diagnostics
into proofs. These boundaries should remain explicit in future releases.

\section{Mathematical boundary}

MathDevMCP can route obligations, expose missing assumptions, run diagnostic
checks, and call deterministic backends when configured. It cannot turn a
natural-language theorem, an incomplete proof sketch, or a nearby equation into
a certified theorem by itself. If a result says \texttt{unverified} or
\texttt{diagnostic\_only}, an agent must preserve that language.

\section{Parser boundary}

Parser evidence is provenance evidence. It tells the user what label, file,
environment, and line span were found. It does not guarantee that every macro,
include, or semantic dependency has been understood. This is why parser policy
and benchmark gates are release artifacts rather than implementation details.

\section{Code-comparison boundary}

Code--document comparison is a drift detector. A missing term can identify a
review target, and a matched term can rule out some stale-code failures. Neither
one is a proof of semantic equivalence. When a project uses domain-specific
idioms such as factorizations, automatic differentiation, or generated kernels,
the benchmark should grow to recognize those idioms explicitly.

\section{Operational boundary}

Full release readiness depends on the operator's environment. LaTeXML, Lean,
LeanDojo, and private-corpus manifests are external capabilities. MathDevMCP
reports them; it does not pretend they exist when they do not. That honesty is
part of the release claim.

\part{Detailed Release Manual}

\chapter{End-to-End Colleague Scenario}

This chapter walks through a realistic review day. A colleague has a document
claim about a state-space likelihood and a Python implementation that is being
edited quickly before a deadline. The colleague does not want a polished essay;
they want to know where the claim lives, whether the code visibly contains the
required numerical operations, whether a derivation can be certified, and what
should be reviewed by a human.

\section{Step 1: locate the candidate claim}

The agent begins with \texttt{search-latex}. This avoids relying on a language
model's memory of the repository. The search result gives labels and local
context. The colleague can inspect those labels directly or ask the agent to use
the top result. If no label is found, the workflow stops as \emph{inconclusive}
rather than inventing a source.

\section{Step 2: extract neighborhood context}

The label neighborhood gives surrounding prose and assumptions. This step
matters because many mathematical claims are not self-contained. A likelihood
equation may rely on notation introduced one paragraph above, while a stability
condition may depend on a boundary condition or discretization convention.

\section{Step 3: compare to code}

The comparison step asks for concrete required terms such as
\texttt{logdet}, \texttt{solve}, \texttt{gradient}, or \texttt{stability}. A
missing term is not automatically a proof of a bug, but it is a precise review
signal. The output includes the selected label, findings, and status, so the
agent can explain what was checked.

\section{Step 4: audit the derivation boundary}

The v2 audit report is where MathDevMCP becomes most valuable for disciplined
agent work. It records parser policy, typed diagnostics, shape diagnostics,
numeric diagnostic suggestions, backend attempts, and a verification boundary.
That boundary text is essential: it tells the colleague whether the output is a
certificate, a diagnostic, or an abstention.

\section{Step 5: produce an implementation brief}

For handoff, the implementation brief bundles the key evidence. A teammate can
read one object rather than rerun every command. This is especially useful when
the agent is asked to create a code-review note, a pull-request comment, or a
maintenance ticket.

\chapter{Release Gate Interpretation}

The release gate should be read as a contract, not a vibe check. A green base
profile means public fixtures, governance, release corpus metadata, parser
policy, and benchmark cases pass. It does not mean every optional backend or
private corpus is available. A green full profile means the strict optional
evidence has also been supplied. This separation lets the team use MathDevMCP
productively on ordinary machines while still preserving a stricter bar for
industrial release.

\section{Blockers}

Blockers represent release conditions that must be fixed or explicitly scoped
out. Examples include benchmark failure, parser policy failure, governance
validation failure, missing strict LaTeXML evidence in the LaTeXML profile,
missing backend evidence in the backend profile, and missing private manifest
evidence in the private-corpus or full profiles.

\section{Caveats}

Caveats are not ignored. They are tracked because a release may be useful while
still carrying environmental or evidence limitations. A dirty worktree is a
caveat because generated reports should name the actual commit. Missing private
corpus evidence is a low-severity caveat in the base profile, but a blocker in
the private-corpus and full profiles.

\section{Expected abstentions}

Expected abstentions protect the product from false confidence. A benchmark can
pass because the tool correctly abstained in a case where certification would
be inappropriate. This is different from a tolerated failure budget. The
current benchmark policy requires all cases to pass.

\chapter{Private Corpus Operating Model}

Private corpus validation is designed for teams that cannot commit research
documents, code paths, or real local directory structures. The populated
manifest lives outside git. The release validator reads it locally, uses
unredacted paths internally to parse documents, and emits redacted reports.

\section{External sanitized corpus}

For autonomous release validation in this workspace, the execution created an
external sanitized corpus under a temporary directory. It includes six release
domains: DSGE macro-finance, stochastic volatility, SDE/PDE numerics, ML/LLM
objectives, Bayesian ELBO/VI, and computational-physics MCMC. The generated
manifest is not committed. The report includes only redacted summaries.

\section{Real private corpus substitution}

A real deployment should replace the sanitized manifest with an approved
external manifest. The same validator and full profile commands apply. The
operator should verify that every release-gated entry has expected labels,
parser backends, expected abstentions or false-confidence seeds, and external
document roots. If a real private corpus covers fewer domains than the
sanitized release corpus, the release report should say exactly that.

\section{Failure modes}

Important private-corpus failure modes are:

\begin{itemize}
  \item missing manifest path;
  \item invalid JSON;
  \item malformed entry fields;
  \item private path inside the repository checkout;
  \item missing document root or code root;
  \item release-gated entry with no expected labels;
  \item release-gated entry with no parser backend;
  \item parser policy not selected for proof audit.
\end{itemize}

\chapter{Backend Environment Operating Model}

Backend tools are intentionally not all installed into the main Python
environment. This is a release-quality design choice. LeanDojo can bring a
dependency stack that conflicts with unrelated document or ML packages. The
backend env checks therefore prove that the optional backend can be invoked
without requiring the main working environment to import every optional module.

\section{Lean}

Lean is detected as an executable, usually through the user-local
\texttt{elan} proxy. The release pins the intended toolchain through
\texttt{MATHDEVMCP\_LEAN\_TOOLCHAIN} for MathDevMCP subprocesses. This avoids
changing the user's global Lean default.

\section{LeanDojo}

LeanDojo is checked as a Python module inside \texttt{mathdevmcp-backends}. A
plain \texttt{doctor} command in the active environment may report LeanDojo as
unavailable. The backend env doctor is the authoritative check for the backend
profile.

\section{LaTeXML}

LaTeXML is a system executable, not a Python dependency. The strict LaTeXML
profile requires it to preserve expected labels on the benchmark corpus. The
current evidence shows label preservation and source provenance. Environment
recognition differs from the lightweight parser, so parser policy still chooses
the current parser for proof-audit routing.

\chapter{Code Architecture for Maintainers}

The implementation should be maintained by preserving the boundary between
library logic, CLI rendering, MCP wrappers, scripts, and documentation.

\section{Library logic}

Library functions build structured dictionaries with contract metadata. They
should be deterministic for the same input files and environment. New behavior
should be added in library code first and exercised by direct tests.

\section{CLI}

The CLI should remain a thin argument parser. It should call library functions,
print JSON, and return meaningful exit codes. It should not contain hidden
business logic that bypasses the tested library surface.

\section{MCP facade}

The MCP facade should wrap the same library functions. Tool names are part of
the colleague-facing API. If a tool is renamed, keep a compatibility alias or
document a breaking change.

\section{Scripts}

Scripts should orchestrate repeatable workflows. They should set
\texttt{PYTHONPATH} explicitly, avoid shell string evaluation for backend
commands, and preserve redaction. Scripts that create generated evidence should
write to ignored or explicitly requested output directories.

\chapter{Adding a New Domain}

To add a new mathematical domain, follow this release path:

\begin{enumerate}[label=\arabic*.]
  \item Add a small public or sanitized fixture with at least one labeled
        mathematical block.
  \item Add paired code showing the relevant operations or a seeded mismatch.
  \item Add an assumption-manifest and convention fixture if the domain relies
        on shape, sign, measure, unit, or support conventions.
  \item Add a domain-template entry only if the decomposition is reusable and
        can name generated obligations, required assumptions, diagnostic routes,
        failure modes, positive fixtures, negative fixtures, and certification
        boundary text.
  \item Extend AST operation extraction only if the new operation is genuinely
        needed.
  \item Add benchmark cases for parser preservation, AST operation coverage,
        and expected abstention behavior.
  \item Add matrix/operator IR or typed-diagnostic coverage if the domain uses
        noncommutative algebra or shape-sensitive identities.
  \item Add release corpus metadata.
  \item Add proof-packet or negative-evidence examples when the domain will be
        used for colleague-facing review.
  \item Add a report case study and generated evidence snippet if the domain is
        release-facing.
  \item Update private manifest templates only with placeholder paths.
\end{enumerate}

This path keeps new domains measurable. It also prevents the product from
accumulating impressive-sounding workflows that do not have regression tests.

\chapter{Adding a New Backend}

A new backend should enter the system as optional evidence. The correct path is
to add a doctor capability, an isolated invocation strategy if needed, timeout
protection, a clear result contract, and tests for unavailable, timeout,
inconclusive, mismatch, and certifying behavior. Only after that should the
backend be used in proof-audit routing or release profiles.

\section{Backend result discipline}

Backend results should distinguish:

\begin{itemize}
  \item backend unavailable;
  \item input not encodable;
  \item timeout;
  \item diagnostic-only evidence;
  \item counterexample or mismatch;
  \item deterministic certificate.
\end{itemize}

The last item is intentionally rare. Most mathematical development tools are
useful because they reduce search space and expose gaps, not because they
automatically certify every claim.

\chapter{Release Evidence Maintenance}

The evidence snippets in this report are generated by
\path{scripts/generate_release_report_evidence.sh}. Regenerate them after
changing release policy, parser policy, benchmark cases, private corpus
validation, or user-facing workflow examples. Do not hand-edit generated
evidence unless the file is explicitly converted into narrative documentation.

\section{Evidence hygiene}

After regeneration, run a path leak scan. The scan should check for real private
roots, local manifest filenames, and repository-local absolute paths. The
committed snippets may contain \texttt{<repo>},
\texttt{<redacted-private-path>}, and
\texttt{<redacted-private-manifest>}.

\section{Evidence reproducibility}

Generated snippets should name the command that produced them. A colleague
reading the report should be able to rerun the command in the repository and
understand why the local result may differ if the worktree is dirty, optional
backends are missing, or a private manifest is not configured.

\chapter{Risk Register}

\section{Risk: overclaiming certification}

The most important product risk is false confidence. Mitigation: keep
verification-boundary text in proof-audit reports, preserve diagnostic-only
flags, and require benchmark cases where abstention is the correct outcome.

\section{Risk: private data leakage}

Private roots and manifests are sensitive. Mitigation: keep populated manifests
outside git, redact normal reports, reject private paths inside the checkout,
and scan generated docs before commit.

\section{Risk: backend dependency conflict}

Optional backends can conflict with the main environment. Mitigation: use
\path{mathdevmcp-backends}, backend-specific doctor checks, and explicit
environment variables rather than broad installation into the base env.

\section{Risk: benchmark drift}

Benchmark totals and category meanings can drift as the project grows.
Mitigation: update tests, reset memo, and release report together whenever a
benchmark case is added or removed.

\section{Risk: large-module entropy}

The codebase has several large modules. Mitigation: keep public facades stable,
extract pure helper modules when it improves locality, and run focused tests
after each extraction.

\appendix
\part*{Appendices}
\addcontentsline{toc}{part}{Appendices}

\chapter{Release Corpus Summary}

\lstinputlisting[caption={Release corpus validation summary},basicstyle=\ttfamily\footnotesize]{generated/release_report/release-corpus-summary.txt}

\chapter{Machine-Readable Evidence Excerpts}

The full generated JSON excerpts remain committed under
\path{docs/generated/release_report/}. They are intentionally referenced here
rather than printed in full so that the report foregrounds colleague workflows
and MCP conversations while preserving reproducibility.

\begin{longtable}{p{0.36\textwidth}p{0.54\textwidth}}
\toprule
Evidence file & Purpose \\
\midrule
\path{benchmark-gate-json-excerpt.txt} & CI-friendly benchmark gate status,
category totals, and expected abstentions. \\
\path{parser-policy-json-excerpt.txt} & Parser role decision and proof-audit
routing boundary. \\
\path{release-corpus-json-excerpt.txt} & Public and configured private corpus
manifest summary with redaction metadata. \\
\path{release-readiness-full-json-excerpt.txt} & Full-profile blockers,
caveats, capability evidence, and profile requirements. \\
\path{private-corpus-json-excerpt.txt} & Private-corpus validation payload with
private paths redacted. \\
\bottomrule
\end{longtable}

\chapter{Evidence Index}

\lstinputlisting[caption={Generated evidence index},basicstyle=\ttfamily\footnotesize]{generated/release_report/evidence-index.txt}

\chapter{Command Reference}

\begin{longtable}{p{0.32\textwidth}p{0.58\textwidth}}
\toprule
Command & Purpose \\
\midrule
\texttt{doctor} & Report runtime capabilities and optional backend status. \\
\texttt{status-taxonomy} & Print the public status and substatus taxonomy. \\
\texttt{search-latex} & Search a LaTeX source tree for relevant context. \\
\texttt{index-latex} & Build a LaTeX index with labels, blocks, and equation-row localization. \\
\texttt{latex-label-lookup} & Fetch a labeled block with paragraph context and provenance. \\
\texttt{audit-implementation-label} & Audit labeled document terms and operations against code. \\
\texttt{domain-templates} & Print the governed diagnostic domain-template catalog. \\
\texttt{suggest-domain-templates} & Suggest domain templates from label, section path, or equation text. \\
\texttt{generate-template-obligations} & Generate diagnostic obligations from a domain template. \\
\texttt{typed-obligation-label} & Build typed and dimensional diagnostics for a labeled obligation. \\
\texttt{check-equality} & Check a bounded equality under explicit assumptions. \\
\texttt{audit-derivation-v2-label} & Build typed proof-audit evidence for a label. \\
\texttt{proof-packet-label} & Build a durable proof packet around a labeled block. \\
\texttt{negative-evidence-label} & Build a packet for blocked, mismatching, or otherwise negative evidence. \\
\texttt{claim-support} & Build a local claim-support packet without treating citations as proof. \\
\texttt{implementation-brief} & Bundle search, context, and code checks. \\
\texttt{lean-readiness} & Report direct Lean, Lake, and LeanDojo readiness separately. \\
\texttt{benchmark-gate} & Run the release benchmark gate. \\
\texttt{release-profile-analysis} & Analyze current readiness across all release profiles. \\
\texttt{validate-release-corpus} & Validate public and configured private corpus metadata. \\
\texttt{release-readiness} & Produce a profile-specific release report. \\
\bottomrule
\end{longtable}

\chapter{Private Manifest Fields}

\begin{longtable}{p{0.32\textwidth}p{0.58\textwidth}}
\toprule
Field & Meaning \\
\midrule
\texttt{id} & Stable entry identifier. \\
\texttt{domain} & Mathematical or colleague workflow domain. \\
\texttt{privacy\_class} & \texttt{private\_external} or \texttt{private\_sanitized\_external}. \\
\texttt{document\_root} & External LaTeX/document root. \\
\texttt{code\_roots} & External code roots associated with the documents. \\
\texttt{expected\_labels} & Labels the parser must preserve. \\
\texttt{expected\_operations} & Mathematical operations expected in code or AST evidence. \\
\texttt{expected\_abstentions} & Known cases where conservative abstention is expected. \\
\texttt{seeded\_false\_confidence\_cases} & False-confidence seeds for review. \\
\texttt{required\_parser\_backends} & Parser backends required for the entry. \\
\texttt{release\_gate\_enabled} & Whether the entry is mandatory for private profile evidence. \\
\texttt{notes} & Human-readable release note. \\
\bottomrule
\end{longtable}

\backmatter
\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
